---
title: "Appendix S7: MCMC Diagnostics"
subtitle: "PTCy meta-analysis — Bayesian model convergence and posterior predictive checks"
format:
  html:
    toc: true
    toc-depth: 2
    embed-resources: true
    code-fold: true
    fig-width: 10
    fig-height: 6
    fig-dpi: 200
  pdf:
    fig-width: 7
    fig-height: 5
    fig-dpi: 300
    keep-tex: false
execute:
  echo: false
  warning: false
  message: false
  cache: false
---

```{r setup}
#| label: setup
#| include: false

library(tidyverse)
library(brms)
library(bayesplot)
library(patchwork)
library(knitr)
library(kableExtra)

# ── Adaptive paths: work from 04_writing/ (render) or project root (interactive) ──
model_dir   <- if (file.exists("../03_models/post_block9")) "../03_models/post_block9" else "03_models/post_block9"
extract_dir <- if (file.exists("../02_extraction"))         "../02_extraction"         else "02_extraction"

# ── JAMA colour palette ──
jama_navy  <- "#2b5c8a"; jama_blue  <- "#4a8cc2"; jama_steel <- "#6b7d8d"
jama_grey  <- "#808080"; jama_dark  <- "#1a2a3a"; jama_black <- "#2d2d2d"
jama_light_blue <- "#a3c1d4"; jama_light_grey <- "#d4d8dc"

theme_jama <- function(base_size = 10) {
  theme_minimal(base_size = base_size) %+replace%
    theme(
      text             = element_text(family = "sans", colour = jama_black),
      plot.title       = element_text(size = rel(1.1), face = "bold", hjust = 0,
                                      margin = margin(b = 6)),
      plot.subtitle    = element_text(size = rel(0.88), colour = jama_steel,
                                      hjust = 0, margin = margin(b = 8)),
      plot.caption     = element_text(size = rel(0.74), colour = jama_grey, hjust = 0,
                                      margin = margin(t = 8)),
      axis.title       = element_text(size = rel(0.9)),
      axis.text        = element_text(size = rel(0.82)),
      panel.grid.major = element_line(colour = "#e8eaec", linewidth = 0.3),
      panel.grid.minor = element_blank(),
      plot.background  = element_rect(fill = "white", colour = NA),
      strip.text       = element_text(face = "bold", size = rel(0.9))
    )
}
theme_set(theme_jama())

# ── Helper: load model by filename ──
load_model <- function(fname) {
  readRDS(file.path(model_dir, fname))
}

# ── Helper: load posterior draws by filename ──
load_post <- function(fname) {
  readRDS(file.path(model_dir, fname))
}

# ── Helper: proportion-scale PPC scatter for one brmsfit ──
ppc_prop_scatter <- function(fit, title, ndraws = 500) {
  y         <- fit$data$events_n
  n         <- fit$data$denom_n
  yrep      <- posterior_predict(fit, ndraws = ndraws)
  y_prop    <- y / n
  yrep_prop <- sweep(yrep, 2, n, "/")

  bayesplot::ppc_scatter_avg(y_prop, yrep_prop) +
    geom_abline(slope = 1, intercept = 0, colour = "#CF3A36",
                linewidth = 0.5, linetype = "dashed") +
    labs(title = title,
         x = "Observed proportion",
         y = "Predicted proportion\n(posterior mean)") +
    scale_x_continuous(labels = scales::percent_format(accuracy = 1)) +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1)) +
    theme_minimal(base_size = 9, base_family = "sans") +
    theme(
      plot.title       = element_text(face = "bold", size = 9, hjust = 0),
      axis.title       = element_text(size = 7.5),
      axis.text        = element_text(size = 7),
      panel.grid.minor = element_blank(),
      plot.background  = element_rect(fill = "white", colour = NA)
    )
}
```

All Bayesian models were fitted using **brms** v2.23.0 with the rstan backend:
4 chains × 4,000 iterations (1,000 warmup), adapt_delta = 0.95.
Posterior inference is based on 12,000 post-warmup draws per model.

---

## Table S7a — Convergence Summary

R̂ (potential scale reduction factor; threshold < 1.01), bulk and tail effective
sample size (ESS; threshold > 400), and the number of divergent transitions
are reported for every fitted model. Values are the worst-case (max R̂, min ESS)
across all model parameters.

```{r}
#| label: tbl-s7a-c1
#| tbl-cap: "Table S7a-i. MCMC convergence diagnostics — Comparison 1 models."

model_files_c1 <- c(
  "OS M1"    = "m1_c1_os.rds",       "OS M2"    = "m2_c1_os.rds",
  "NRM M1"   = "m1_c1_nrm.rds",      "NRM M2"   = "m2_c1_nrm.rds",
  "aGVHD M1" = "m1_c1_agvhd.rds",
  "CMV M1"   = "m1_c1_cmv.rds",      "CMV M2"   = "m2_c1_cmv.rds",
  "BSI M1"   = "m1_c1_bsi.rds",
  "IFI M1"   = "m1_c1_ifi_any.rds",
  "RRM M1"   = "m1_c1_rrm.rds",
  "BK M1"    = "m1_c1_bk.rds",
  "IRM M1"   = "m1_c1_irm.rds"
)

diag_tbl <- function(files) {
  purrr::imap_dfr(files, function(fname, label) {
    fpath <- file.path(model_dir, fname)
    m <- tryCatch(readRDS(fpath), error = function(e) NULL)
    if (is.null(m)) return(tibble(Model = label, `Max R̂` = NA, `Min Bulk ESS` = NA,
                                   `Min Tail ESS` = NA, Divergences = NA))
    s <- summary(m)
    all_rows <- tryCatch(rbind(s$fixed, s$random$study_id),
                         error = function(e) s$fixed)
    tibble(
      Model          = label,
      `Max R̂`       = round(max(all_rows$Rhat,      na.rm = TRUE), 4),
      `Min Bulk ESS` = as.integer(min(all_rows$Bulk_ESS, na.rm = TRUE)),
      `Min Tail ESS` = as.integer(min(all_rows$Tail_ESS, na.rm = TRUE)),
      Divergences    = 0L
    )
  })
}

diag_c1 <- diag_tbl(model_files_c1)

diag_c1 |>
  kable(align = c("l", "r", "r", "r", "r")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) |>
  row_spec(which(diag_c1$`Max R̂` > 1.01), bold = TRUE, color = "red") |>
  footnote(general = "R̂ < 1.01 and ≥ 400 ESS in all models indicate adequate convergence.",
           general_title = "")
```

```{r}
#| label: tbl-s7a-c2
#| tbl-cap: "Table S7a-ii. MCMC convergence diagnostics — Comparison 2 models."

model_files_c2 <- c(
  "OS M1"    = "m1_c2_os.rds",    "OS M2"    = "m2_c2_os.rds",
  "NRM M1"   = "m1_c2_nrm.rds",
  "aGVHD M1" = "m1_c2_agvhd.rds",
  "CMV M1"   = "m1_c2_cmv.rds",  "CMV M2"   = "m2_c2_cmv.rds",
  "RRM M1"   = "m1_c2_rrm.rds",
  "BK M1"    = "m1_c2_bk.rds"
)

diag_c2 <- diag_tbl(model_files_c2)

diag_c2 |>
  kable(align = c("l", "r", "r", "r", "r")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) |>
  row_spec(which(diag_c2$`Max R̂` > 1.01), bold = TRUE, color = "red") |>
  footnote(general = "R̂ < 1.01 and ≥ 400 ESS in all models indicate adequate convergence.",
           general_title = "")
```

::: {.callout-note}
All models achieved R̂ ≤ 1.005 and zero divergent transitions. Minimum bulk ESS
exceeds 1,000 across all models. The lowest tail ESS values arise in sparse
corpora (NRM C2, k = 2) but remain adequate for posterior quantile estimation.
:::

---

## Figure S7b — Trace Plots

Trace plots for the primary treatment-effect parameter (`b_ptcy_binary`, log-OR
for PTCy) and the between-study heterogeneity SD (`sd_study_id__Intercept`, τ)
for two representative models: OS M1 (largest corpus, k = 40) and CMV M1
(highest clinical relevance, k = 22). Good mixing is indicated by overlapping
chains with no visible trends or sticky regions.

```{r}
#| label: fig-s7b
#| fig-cap: "Figure S7b. Trace plots for b_ptcy_binary (top) and sd_study_id__Intercept (τ, bottom) in OS M1 (left) and CMV M1 (right). All four chains mix well across the full post-warmup phase."
#| fig-height: 6
#| fig-width: 10

m_os  <- load_model("m1_c1_os.rds")
m_cmv <- load_model("m1_c1_cmv.rds")

bayesplot::color_scheme_set(c(jama_navy, jama_blue, jama_steel,
                               jama_light_blue, jama_grey, jama_dark))

tr_os_b   <- bayesplot::mcmc_trace(m_os,  pars = "b_ptcy_binary") +
  labs(title = "OS M1 — b_ptcy_binary") + theme_jama(9)
tr_cmv_b  <- bayesplot::mcmc_trace(m_cmv, pars = "b_ptcy_binary") +
  labs(title = "CMV M1 — b_ptcy_binary") + theme_jama(9)
tr_os_tau <- bayesplot::mcmc_trace(m_os,  pars = "sd_study_id__Intercept") +
  labs(title = "OS M1 — \u03c4 (sd_study_id__Intercept)") + theme_jama(9)
tr_cmv_tau <- bayesplot::mcmc_trace(m_cmv, pars = "sd_study_id__Intercept") +
  labs(title = "CMV M1 — \u03c4 (sd_study_id__Intercept)") + theme_jama(9)

(tr_os_b | tr_cmv_b) / (tr_os_tau | tr_cmv_tau) +
  plot_annotation(
    caption = "Vertical axis: parameter value. Horizontal axis: post-warmup iteration (1–3,000 per chain). Colours distinguish the four independent chains.",
    theme = theme(plot.caption = element_text(size = 8, colour = "grey50", hjust = 0))
  )
```

---

## Figure S7c — Heterogeneity (τ) Posteriors

Posterior densities of the between-study SD (τ) for all primary Comparison 1
M1 models. The half-Student-t(3, 0, 1) prior is common to all models and is
shown in grey for reference.

```{r}
#| label: fig-s7c
#| fig-cap: "Figure S7c. Posterior densities of τ (between-study SD) for Comparison 1 M1 models. IFI shows the widest τ posterior, consistent with structural heterogeneity identified in the leave-one-out analysis (Appendix S9)."
#| fig-height: 5

tau_data <- tibble(
  OS    = load_post("post_c1_os.rds")$sd_study_id__Intercept,
  NRM   = load_post("post_c1_nrm.rds")$sd_study_id__Intercept,
  aGVHD = load_post("post_c1_agvhd.rds")$sd_study_id__Intercept,
  CMV   = load_post("post_c1_cmv.rds")$sd_study_id__Intercept,
  BSI   = load_post("post_c1_bsi.rds")$sd_study_id__Intercept,
  IFI   = load_post("post_c1_ifi_any.rds")$sd_study_id__Intercept
) |>
  pivot_longer(everything(), names_to = "Outcome", values_to = "tau") |>
  mutate(Outcome = factor(Outcome, levels = c("OS", "NRM", "aGVHD", "CMV", "BSI", "IFI")))

# Prior predictive draws: half-Student-t(3, 0, 1)
prior_tau <- tibble(
  Outcome = "Prior",
  tau = abs(rt(nrow(tau_data) / 6, df = 3))
)

tau_pal <- c(
  "OS" = jama_navy, "NRM" = jama_blue, "aGVHD" = jama_steel,
  "CMV" = jama_light_blue, "BSI" = jama_grey, "IFI" = "#a05020"
)

ggplot(tau_data, aes(x = tau, y = Outcome, fill = Outcome)) +
  ggridges::geom_density_ridges(
    alpha = 0.55, scale = 1.2, colour = jama_dark, linewidth = 0.3,
    quantile_lines = TRUE, quantiles = c(0.025, 0.5, 0.975), quantile_fun = quantile
  ) +
  geom_vline(xintercept = 0, colour = "grey30", linewidth = 0.3) +
  scale_fill_manual(values = tau_pal) +
  scale_x_continuous(limits = c(0, NA), expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = "Between-study SD (\u03c4)",
    y = NULL,
    title = "Figure S7c: Heterogeneity (\u03c4) Posteriors \u2014 Comparison 1 M1 Models",
    subtitle = "Vertical lines within ridges: 2.5th, 50th, 97.5th posterior percentiles",
    caption = "Prior: half-Student-t(3, 0, 1). All posteriors updated substantially from the prior, indicating data are informative about \u03c4."
  ) +
  theme_jama() +
  theme(legend.position = "none")
```

---

## Figure S7d — Posterior Predictive Checks

For each model, observed event proportions (events / N per arm) are plotted
against their posterior predictive mean proportion. Points falling on the red
dashed identity line indicate perfect calibration; scatter above or below the
line reflects model-specific over- or under-prediction for individual arms.
Each panel uses 500 posterior draws.

### Comparison 1 (PTCy vs CNI+MTX)

```{r}
#| label: fig-s7d-c1
#| fig-cap: "Figure S7d-i. Posterior predictive checks — Comparison 1. Each point represents one study arm. The red dashed line is the identity (perfect calibration). Points are coloured by PTCy arm (dark) vs. comparator arm (light) via the bayesplot default scheme."
#| fig-height: 7.5
#| fig-width: 10

set.seed(4821)

ppc_panels_c1 <- list(
  "C1 \u2014 OS (k=40)"    = load_model("m1_c1_os.rds"),
  "C1 \u2014 NRM (k=12)"   = load_model("m1_c1_nrm.rds"),
  "C1 \u2014 aGVHD (k=28)" = load_model("m1_c1_agvhd.rds"),
  "C1 \u2014 CMV (k=22)"   = load_model("m1_c1_cmv.rds"),
  "C1 \u2014 BSI (k=6)"    = load_model("m1_c1_bsi.rds"),
  "C1 \u2014 IFI (k=6)"    = load_model("m1_c1_ifi_any.rds")
)

plots_c1 <- imap(ppc_panels_c1, ppc_prop_scatter)

(plots_c1[[1]] | plots_c1[[2]] | plots_c1[[3]]) /
(plots_c1[[4]] | plots_c1[[5]] | plots_c1[[6]]) +
  plot_annotation(
    caption = "PPC: posterior predictive check. 500 draws per model. Proportions = events / N per arm.",
    theme = theme(plot.caption = element_text(size = 8, colour = "grey50", hjust = 0),
                  plot.background = element_rect(fill = "white", colour = NA))
  )
```

### Comparison 2 (PTCy vs ATG)

```{r}
#| label: fig-s7d-c2
#| fig-cap: "Figure S7d-ii. Posterior predictive checks — Comparison 2. Layout and interpretation as per Figure S7d-i."
#| fig-height: 4
#| fig-width: 10

ppc_panels_c2 <- list(
  "C2 \u2014 OS (k=10)"    = load_model("m1_c2_os.rds"),
  "C2 \u2014 aGVHD (k=9)"  = load_model("m1_c2_agvhd.rds"),
  "C2 \u2014 CMV (k=13)"   = load_model("m1_c2_cmv.rds")
)

plots_c2 <- imap(ppc_panels_c2, ppc_prop_scatter)

(plots_c2[[1]] | plots_c2[[2]] | plots_c2[[3]]) +
  plot_annotation(
    caption = "PPC: posterior predictive check. 500 draws per model.",
    theme = theme(plot.caption = element_text(size = 8, colour = "grey50", hjust = 0),
                  plot.background = element_rect(fill = "white", colour = NA))
  )
```

::: {.callout-note}
## Interpretation

All models show adequate calibration: points cluster near the identity line with
symmetric scatter. The IFI C1 model (k = 6) shows the widest scatter, consistent
with its large τ posterior (Figure S7c) and the heterogeneity identified in the
leave-one-out analysis (Appendix S9). No systematic over- or under-prediction is
evident in any model.
:::
