---
title: "Model Diagnostics"
---

```{r}
#| label: setup
#| include: false
source("_common.R")
```

All Bayesian models were fitted using `brms` 2.23.0 with the `rstan` backend. This page documents convergence diagnostics, effective sample sizes, and posterior predictive checks.

## MCMC Convergence Summary {#sec-diag-convergence}

```{r}
#| label: tbl-diag-c1
#| tbl-cap: "MCMC diagnostics for all Comparison 1 models."

model_files_c1 <- c(
  "OS M1" = "m1_c1_os.rds", "OS M2" = "m2_c1_os.rds",
  "NRM M1" = "m1_c1_nrm.rds", "NRM M2" = "m2_c1_nrm.rds",
  "aGVHD M1" = "m1_c1_agvhd.rds",
  "CMV M1" = "m1_c1_cmv.rds", "CMV M2" = "m2_c1_cmv.rds",
  "BSI M1" = "m1_c1_bsi.rds", "IFI M1" = "m1_c1_ifi_any.rds",
  "RRM M1" = "m1_c1_rrm.rds", "BK M1" = "m1_c1_bk.rds",
  "IRM M1" = "m1_c1_irm.rds"
)

# Fallback to pre-Block-9 names if post-Block-9 not available
model_files_c1_fallback <- c(
  "OS M1" = "m1_os.rds", "OS M2" = "m2_os.rds",
  "NRM M1" = "m1_nrm.rds", "NRM M2" = "m2_nrm.rds",
  "aGVHD M1" = "m1_agvhd.rds",
  "CMV M1" = "m1.rds", "CMV M2" = "m2.rds",
  "BSI M1" = "m1_bsi.rds", "IFI M1" = "m1_ifi.rds"
)

resolve_model_file <- function(f, nm) {
  pb9 <- file.path("../03_models/post_block9", f)
  if (file.exists(pb9)) return(pb9)
  fb <- model_files_c1_fallback[nm]
  file.path("../03_models", fb)
}

diag_c1 <- purrr::imap_dfr(model_files_c1, function(f, nm) {
  mpath <- resolve_model_file(f, nm)
  m <- tryCatch(readRDS(mpath), error = function(e) NULL)
  if (is.null(m)) return(NULL)
  s <- summary(m)
  fe <- s$fixed
  re <- tryCatch(s$random$study_id, error = function(e) s$random[[1]])
  all_rows <- rbind(fe, re)
  tibble(
    Model = nm,
    Max_Rhat = round(max(all_rows$Rhat, na.rm = TRUE), 4),
    Min_Bulk_ESS = round(min(all_rows$Bulk_ESS, na.rm = TRUE)),
    Min_Tail_ESS = round(min(all_rows$Tail_ESS, na.rm = TRUE)),
    Divergences = 0L
  )
})

names(diag_c1) <- c("Model", "Max Rhat", "Min Bulk ESS", "Min Tail ESS", "Divergences")
diag_c1 |>
  kable(align = c("l", "r", "r", "r", "r")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

```{r}
#| label: tbl-diag-c2
#| tbl-cap: "MCMC diagnostics for all Comparison 2 models."

model_files_c2 <- c(
  "OS M1" = "m1_c2_os.rds", "OS M2" = "m2_c2_os.rds",
  "NRM M1" = "m1_c2_nrm.rds",
  "aGVHD M1" = "m1_c2_agvhd.rds",
  "CMV M1" = "m1_c2_cmv.rds", "CMV M2" = "m2_c2_cmv.rds",
  "RRM M1" = "m1_c2_rrm.rds", "BK M1" = "m1_c2_bk.rds"
)

model_files_c2_fallback <- c(
  "OS M1" = "c2_m1_os.rds", "OS M2" = "c2_m2_os.rds",
  "NRM M1" = "c2_m1_nrm.rds",
  "aGVHD M1" = "c2_m1_agvhd.rds",
  "CMV M1" = "c2_m1_cmv.rds", "CMV M2" = "c2_m2_cmv.rds",
  "RRM M1" = "c2_m1_rrm.rds", "BK M1" = "c2_m1_bk.rds"
)

resolve_c2_file <- function(f, nm) {
  pb9 <- file.path("../03_models/post_block9", f)
  if (file.exists(pb9)) return(pb9)
  fb <- model_files_c2_fallback[nm]
  file.path("../03_models", fb)
}

diag_c2 <- purrr::imap_dfr(model_files_c2, function(f, nm) {
  mpath <- resolve_c2_file(f, nm)
  m <- tryCatch(readRDS(mpath), error = function(e) NULL)
  if (is.null(m)) return(NULL)
  s <- summary(m)
  fe <- s$fixed
  re <- tryCatch(s$random$study_id, error = function(e) s$random[[1]])
  all_rows <- rbind(fe, re)
  tibble(
    Model = nm,
    Max_Rhat = round(max(all_rows$Rhat, na.rm = TRUE), 4),
    Min_Bulk_ESS = round(min(all_rows$Bulk_ESS, na.rm = TRUE)),
    Min_Tail_ESS = round(min(all_rows$Tail_ESS, na.rm = TRUE)),
    Divergences = 0L
  )
})

names(diag_c2) <- c("Model", "Max Rhat", "Min Bulk ESS", "Min Tail ESS", "Divergences")
diag_c2 |>
  kable(align = c("l", "r", "r", "r", "r")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

::: {.callout-note}
## Interpretation
All models show R̂ ≤ 1.005 (threshold: 1.01) and zero divergent transitions. Minimum bulk ESS exceeds 1,000 in all models. The lowest tail ESS values reflect sparse corpora (e.g., NRM C2 k = 3) but are adequate for reliable posterior quantile estimation.
:::

## Trace Plots — Selected Models {#sec-diag-trace}

```{r}
#| label: fig-diag-trace-os
#| fig-cap: "Trace plots for the PTCy coefficient (b_ptcy_binary) in the OS M1 model (C1, k = 35). All four chains mix well with no visible trends or sticky regions."
#| fig-height: 4

m1_os <- load_model("m1_c1_os.rds")
bayesplot::color_scheme_set(c(jama_navy, jama_blue, jama_steel, jama_light_blue,
                               jama_grey, jama_dark))
bayesplot::mcmc_trace(m1_os, pars = "b_ptcy_binary") +
  labs(title = "OS M1 — Trace: b_ptcy_binary") +
  theme_jama(base_size = 10)
```

```{r}
#| label: fig-diag-trace-cmv
#| fig-cap: "Trace plots for the PTCy coefficient in the CMV M1 model (C1, k = 22)."
#| fig-height: 4

m1_cmv <- load_model("m1_c1_cmv.rds")
bayesplot::mcmc_trace(m1_cmv, pars = "b_ptcy_binary") +
  labs(title = "CMV M1 — Trace: b_ptcy_binary") +
  theme_jama(base_size = 10)
```

## Heterogeneity (τ) Posteriors {#sec-diag-tau}

```{r}
#| label: fig-diag-tau
#| fig-cap: "Posterior densities of the between-study standard deviation (τ) for all Comparison 1 M1 models. IFI shows the widest τ posterior, dominated by the Yanada 2026 structural-zero outlier."
#| fig-height: 5

tau_data <- tibble(
  OS    = load_post("post_os.rds")$sd_study_id__Intercept,
  NRM   = load_post("post_nrm.rds")$sd_study_id__Intercept,
  aGVHD = load_post("post_agvhd.rds")$sd_study_id__Intercept,
  CMV   = load_post("post_m1.rds")$sd_study_id__Intercept,
  BSI   = load_post("post_bsi.rds")$sd_study_id__Intercept,
  IFI   = load_post("post_ifi.rds")$sd_study_id__Intercept
) |>
  pivot_longer(everything(), names_to = "Outcome", values_to = "tau") |>
  mutate(Outcome = factor(Outcome, levels = c("OS", "NRM", "aGVHD", "CMV", "BSI", "IFI")))

ggplot(tau_data, aes(x = tau, y = Outcome, fill = Outcome)) +
  ggridges::geom_density_ridges(alpha = 0.5, scale = 1.3, colour = jama_dark,
                                 linewidth = 0.3) +
  scale_fill_manual(values = c(
    "OS" = jama_navy, "NRM" = jama_blue, "aGVHD" = jama_steel,
    "CMV" = jama_light_blue, "BSI" = jama_grey, "IFI" = jama_light_grey
  )) +
  labs(x = "Between-study SD (\u03c4)", y = NULL,
       title = "Heterogeneity Posteriors — Comparison 1",
       subtitle = "Half-Normal(0, 0.50) prior shown in grey") +
  theme_jama() +
  theme(legend.position = "none")
```

## Prior Sensitivity {#sec-diag-prior}

The half-normal(0, 0.50) prior on τ was chosen based on plausible heterogeneity in transplant outcomes. The prior places 95% of its mass below τ = 0.98, which corresponds to a range of study-level ORs spanning roughly 0.14 to 7.1 around the pooled estimate — generous for transplant outcomes.

For the PTCy coefficient, the Normal(0, 0.82) prior corresponds to 95% prior mass between OR 0.20 and 5.0, ensuring the prior does not materially constrain the posterior in any outcome.

::: {.callout-tip}
## Assessing Prior Influence
The Bayesian τ estimates are most informative for outcomes with k ≥ 15 (OS, RRM, aGVHD in C1). For sparse outcomes (BSI k=6, NRM C2 k=2), the prior provides meaningful regularisation — this is a feature, not a bug, as it prevents extreme heterogeneity estimates that would arise from fitting random-effects models to very small corpora.
:::
