---
title: "Discussion"
---

```{r}
#| label: setup
#| include: false
source("_common.R")

post_os     <- load_post("post_os.rds")
post_m1     <- load_post("post_m1.rds")
post_c2_cmv <- load_post("post_c2_cmv.rds")
```

## The Mechanistic Asymmetry {#sec-disc-asymmetry}

The totality of findings across six pre-specified outcomes and two independent comparisons resolves into a single, internally consistent mechanistic framework built on the distinction between two immunological consequences of PTCy.

### Pathway 1: GVHD Suppression (Intended)

PTCy eliminates alloreactive T cells during the critical window of allorecognition, reducing acute GVHD grade II–IV by an estimated OR of 0.63 (C1) and 0.59 (C2). This propagates downstream:

- **NRM** benefit is completely steroid-mediated in C1 (M1 OR 0.86 → M2 OR 1.02) and substantially mediated in C2 (M1 OR 0.65 → M2 OR 0.76)
- **OS** benefit is only partially mediated (C1: 29% attenuation; C2: 35% attenuation), suggesting an additional direct contribution — most plausibly graft-versus-leukaemia preservation

### Pathway 2: T-Cell Lymphodepletion (Unintended)

The same cyclophosphamide pulse depletes the nascent CMV- and bacterial-specific T-cell repertoire. This manifests as elevated infection risk — but the direction depends on the comparator's own depletion depth.

```{r}
#| label: fig-disc-cmv-hierarchy
#| fig-cap: "T-cell depletion depth hierarchy. CNI-based regimens preserve T cells most; ATG depletes most deeply; PTCy falls between. CMV risk tracks this ordering."
#| fig-height: 4
#| fig-width: 8

hierarchy <- tibble(
  Regimen = factor(c("CNI+MTX/MMF", "PTCy", "ATG"),
                    levels = c("CNI+MTX/MMF", "PTCy", "ATG")),
  depletion = c(1, 2, 3),
  cmv_label = c("Lowest CMV\nrisk", "Intermediate", "Highest CMV\nrisk")
)

ggplot(hierarchy, aes(x = Regimen, y = depletion)) +
  geom_col(fill = c(jama_light_blue, jama_navy, jama_dark), width = 0.6) +
  geom_text(aes(label = cmv_label), y = 0.3, colour = "white",
            size = 3.5, fontface = "bold", vjust = 0) +
  annotate("segment", x = 0.5, xend = 3.5, y = 3.3, yend = 3.3,
           arrow = arrow(length = unit(0.2, "cm")), colour = jama_steel) +
  annotate("text", x = 2, y = 3.5, label = "Increasing T-cell depletion depth",
           colour = jama_steel, size = 3.5, fontface = "italic") +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 4)) +
  labs(y = "Relative T-cell depletion", x = NULL,
       title = "T-Cell Depletion Hierarchy and CMV Risk") +
  theme_jama() +
  theme(axis.text.y = element_blank(), axis.ticks.y = element_blank(),
        panel.grid.major.x = element_blank())
```

### The Natural Experiment

Comparison 2 provides a natural test of the framework. If PTCy's CMV excess is driven by depth of depletion (not intrinsic to cyclophosphamide), then comparing PTCy to a deeper depleter should reverse the CMV direction. This is exactly what was observed:

- **C1 (vs CNI+MTX/MMF):** OR 1.41 [1.20–1.65] — PTCy harmful
- **C2 (vs ATG):** OR 0.77 [0.60–0.99] — PTCy protective

```{r}
#| label: fig-disc-overlay
#| fig-cap: "Posterior overlap plot for the CMV effect across comparisons. The distributions are almost entirely non-overlapping, confirming a qualitative direction reversal rather than sampling noise."
#| fig-height: 4

tibble(
  `C1: vs CNI+MTX` = post_m1$b_ptcy_binary,
  `C2: vs ATG` = post_c2_cmv$b_ptcy_binary
) |>
  pivot_longer(everything(), names_to = "Comparison", values_to = "log_or") |>
  ggplot(aes(x = log_or, fill = Comparison, colour = Comparison)) +
  geom_density(alpha = 0.3, linewidth = 0.7) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = jama_grey) +
  scale_fill_manual(values = c("C1: vs CNI+MTX" = jama_navy, "C2: vs ATG" = jama_steel)) +
  scale_colour_manual(values = c("C1: vs CNI+MTX" = jama_navy, "C2: vs ATG" = jama_steel)) +
  labs(x = "log-OR (PTCy vs comparator)", y = "Posterior density",
       title = "CMV Direction Reversal — Posterior Overlap") +
  theme_jama()
```

## Practical Implications {#sec-disc-practical}

The mechanistic architecture carries direct clinical implications:

| Domain | Intervention target | Rationale |
|---|---|---|
| **CMV / BSI** | T-cell reconstitution (adoptive therapy, early IS tapering) | Infection risk is GVHD-independent; steroid avoidance will not help |
| **NRM** | GVHD control and steroid minimisation | NRM benefit is fully steroid-mediated |
| **OS** | Both GVHD control + relapse prevention | Partial mediation suggests dual contribution |

::: {.callout-note}
## ATG Is Not the Solution
The C2 data show that ATG-based regimens carry at least as great a T-cell reconstitution liability as PTCy. The infection costs of PTCy are best understood as a general consequence of T-cell–depleting prophylaxis, not a PTCy-specific limitation.
:::

## Limitations {#sec-disc-limitations}

1. **Observational evidence.** All findings derive from retrospective cohort and registry studies. Residual confounding cannot be excluded despite mediation modelling.
2. **Sparse infection data.** BSI (k = 3–6) and IFI (k = 4–8) are data-poor. Dedicated prospective studies are needed.
3. **Relapse data not analysed.** The residual OS benefit (surviving steroid adjustment) is hypothesised to reflect graft-versus-leukaemia preservation, but relapse data were not extracted.
4. **Publication bias.** Formally assessed for C1 outcomes with k ≥ 10; not assessable for most C2 outcomes.
5. **Extraction incomplete.** Blocks 7–8 (99 papers) remain unextracted; the final corpus may shift some estimates.
