---
title: "Study-Level Forest Plots"
---

```{r}
#| label: setup
#| include: false
source("_common.R")

studies_df <- read_csv(file.path(extract_dir, "studies.csv"), show_col_types = FALSE)

# C1 paired data
c1_os    <- read.csv(file.path(model_dir, "data_c1_os.csv"))
c1_nrm   <- read.csv(file.path(model_dir, "data_c1_nrm.csv"))
c1_agvhd <- read.csv(file.path(model_dir, "data_c1_agvhd.csv"))
c1_cmv   <- read.csv(file.path(model_dir, "data_c1_cmv.csv"))
c1_bsi   <- read.csv(file.path(model_dir, "data_c1_bsi.csv"))
c1_ifi   <- read.csv(file.path(model_dir, "data_c1_ifi_any.csv"))
c1_rrm   <- tryCatch(read.csv(file.path(model_dir, "data_c1_rrm.csv")), error = function(e) NULL)
c1_bk    <- tryCatch(read.csv(file.path(model_dir, "data_c1_bk.csv")), error = function(e) NULL)
c1_irm     <- tryCatch(read.csv(file.path(model_dir, "data_c1_irm.csv")), error = function(e) NULL)
c1_cgvhd_f <- tryCatch(read.csv(file.path(model_dir, "data_c1_cgvhd_ms.csv")), error = function(e) NULL)

# C2 paired data
c2_os    <- read.csv(file.path(model_dir, "data_c2_os.csv"))
c2_nrm   <- read.csv(file.path(model_dir, "data_c2_nrm.csv"))
c2_agvhd <- read.csv(file.path(model_dir, "data_c2_agvhd.csv"))
c2_cmv   <- read.csv(file.path(model_dir, "data_c2_cmv.csv"))
c2_rrm   <- tryCatch(read.csv(file.path(model_dir, "data_c2_rrm.csv")), error = function(e) NULL)
c2_bk    <- tryCatch(read.csv(file.path(model_dir, "data_c2_bk.csv")), error = function(e) NULL)
c2_irm     <- tryCatch(read.csv(file.path(model_dir, "data_c2_irm.csv")), error = function(e) NULL)
c2_cgvhd_f <- tryCatch(read.csv(file.path(model_dir, "data_c2_cgvhd_ms.csv")), error = function(e) NULL)
```

Study-level forest plots show individual study odds ratios with 95% confidence intervals from frequentist random-effects meta-analysis (REML estimator). The pooled estimate is displayed as a diamond at the top. OR \< 1 favours PTCy for all outcomes.

## Comparison 1: PTCy vs CNI+MTX/MMF {#sec-c1}

### Overall Survival

```{r}
#| label: fig-forest-c1-os
#| fig-cap: "Forest plot — Overall survival, C1."
#| fig-height: !expr max(6, nrow(c1_os) * 0.35 + 2)
#| fig-width: 10
forest_study(c1_os, studies_df, "Overall Survival", "C1: PTCy vs CNI+MTX/MMF")
```

### Non-Relapse Mortality

```{r}
#| label: fig-forest-c1-nrm
#| fig-cap: "Forest plot — Non-relapse mortality, C1."
#| fig-height: !expr max(5, nrow(c1_nrm) * 0.35 + 2)
#| fig-width: 10
forest_study(c1_nrm, studies_df, "Non-Relapse Mortality", "C1: PTCy vs CNI+MTX/MMF")
```

### Acute GVHD Grade II–IV

```{r}
#| label: fig-forest-c1-agvhd
#| fig-cap: "Forest plot — Acute GVHD grade II–IV, C1."
#| fig-height: !expr max(5, nrow(c1_agvhd) * 0.35 + 2)
#| fig-width: 10
forest_study(c1_agvhd, studies_df, "Acute GVHD Grade II\u2013IV", "C1: PTCy vs CNI+MTX/MMF")
```

### Chronic GVHD Moderate-Severe

```{r}
#| label: fig-forest-c1-cgvhd
#| fig-cap: "Forest plot — Chronic GVHD moderate-severe, C1."
#| fig-height: 10
#| fig-width: 10

if (!is.null(c1_cgvhd_f)) forest_study(c1_cgvhd_f, studies_df, "Chronic GVHD Moderate-Severe", "C1: PTCy vs CNI+MTX/MMF")
```

### CMV Reactivation

```{r}
#| label: fig-forest-c1-cmv
#| fig-cap: "Forest plot — CMV any reactivation, C1."
#| fig-height: !expr max(5, nrow(c1_cmv) * 0.35 + 2)
#| fig-width: 10
forest_study(c1_cmv, studies_df, "CMV Any Reactivation", "C1: PTCy vs CNI+MTX/MMF")
```

### Bloodstream Infection

```{r}
#| label: fig-forest-c1-bsi
#| fig-cap: "Forest plot — Bloodstream infection, C1."
#| fig-height: !expr max(4, nrow(c1_bsi) * 0.35 + 2)
#| fig-width: 10
forest_study(c1_bsi, studies_df, "Bloodstream Infection", "C1: PTCy vs CNI+MTX/MMF")
```

### Invasive Fungal Infection

```{r}
#| label: fig-forest-c1-ifi
#| fig-cap: "Forest plot — Invasive fungal infection, C1."
#| fig-height: !expr max(4, nrow(c1_ifi) * 0.35 + 2)
#| fig-width: 10
forest_study(c1_ifi, studies_df, "Invasive Fungal Infection", "C1: PTCy vs CNI+MTX/MMF")
```

### Relapse-Related Mortality

```{r}
#| label: fig-forest-c1-rrm
#| fig-cap: "Forest plot — Relapse-related mortality, C1."
#| fig-height: 12
#| fig-width: 10

if (!is.null(c1_rrm)) forest_study(c1_rrm, studies_df, "Relapse-Related Mortality", "C1: PTCy vs CNI+MTX/MMF")
```

### BK Virus Reactivation

```{r}
#| label: fig-forest-c1-bk
#| fig-cap: "Forest plot — BK virus reactivation, C1."
#| fig-height: 6
#| fig-width: 10

if (!is.null(c1_bk)) forest_study(c1_bk, studies_df, "BK Virus Reactivation", "C1: PTCy vs CNI+MTX/MMF")
```

### Infection-Related Mortality

```{r}
#| label: fig-forest-c1-irm
#| fig-cap: "Forest plot — Infection-related mortality, C1."
#| fig-height: 8
#| fig-width: 10

if (!is.null(c1_irm)) forest_study(c1_irm, studies_df, "Infection-Related Mortality", "C1: PTCy vs CNI+MTX/MMF")
```

---

## Comparison 2: PTCy vs ATG {#sec-c2}

### Overall Survival

```{r}
#| label: fig-forest-c2-os
#| fig-cap: "Forest plot — Overall survival, C2."
#| fig-height: !expr max(4, nrow(c2_os) * 0.35 + 2)
#| fig-width: 10
forest_study(c2_os, studies_df, "Overall Survival", "C2: PTCy vs ATG")
```

### Non-Relapse Mortality

```{r}
#| label: fig-forest-c2-nrm
#| fig-cap: "Forest plot — Non-relapse mortality, C2."
#| fig-height: !expr max(4, nrow(c2_nrm) * 0.35 + 2)
#| fig-width: 10
forest_study(c2_nrm, studies_df, "Non-Relapse Mortality", "C2: PTCy vs ATG")
```

### Acute GVHD Grade II–IV

```{r}
#| label: fig-forest-c2-agvhd
#| fig-cap: "Forest plot — Acute GVHD grade II–IV, C2."
#| fig-height: !expr max(4, nrow(c2_agvhd) * 0.35 + 2)
#| fig-width: 10
forest_study(c2_agvhd, studies_df, "Acute GVHD Grade II\u2013IV", "C2: PTCy vs ATG")
```

### Chronic GVHD Moderate-Severe

```{r}
#| label: fig-forest-c2-cgvhd
#| fig-cap: "Forest plot — Chronic GVHD moderate-severe, C2."
#| fig-height: 5
#| fig-width: 10

if (!is.null(c2_cgvhd_f)) forest_study(c2_cgvhd_f, studies_df, "Chronic GVHD Moderate-Severe", "C2: PTCy vs ATG")
```

### CMV Reactivation

```{r}
#| label: fig-forest-c2-cmv
#| fig-cap: "Forest plot — CMV any reactivation, C2."
#| fig-height: !expr max(4, nrow(c2_cmv) * 0.35 + 2)
#| fig-width: 10
forest_study(c2_cmv, studies_df, "CMV Any Reactivation", "C2: PTCy vs ATG")
```

### Relapse-Related Mortality

```{r}
#| label: fig-forest-c2-rrm
#| fig-cap: "Forest plot — Relapse-related mortality, C2."
#| fig-height: 6
#| fig-width: 10

if (!is.null(c2_rrm)) forest_study(c2_rrm, studies_df, "Relapse-Related Mortality", "C2: PTCy vs ATG")
```

### BK Virus Reactivation

```{r}
#| label: fig-forest-c2-bk
#| fig-cap: "Forest plot — BK virus reactivation, C2."
#| fig-height: 5
#| fig-width: 10

if (!is.null(c2_bk)) forest_study(c2_bk, studies_df, "BK Virus Reactivation", "C2: PTCy vs ATG")
```

### Infection-Related Mortality

```{r}
#| label: fig-forest-c2-irm
#| fig-cap: "Forest plot — Infection-related mortality, C2."
#| fig-height: 5
#| fig-width: 10

if (!is.null(c2_irm)) forest_study(c2_irm, studies_df, "Infection-Related Mortality", "C2: PTCy vs ATG")
```
