---
title: "GRADE Certainty Assessment"
---

```{r}
#| label: setup
#| include: false
source("_common.R")
```

GRADE certainty was assessed per the GRADE Handbook (Schünemann et al.), adapted for Bayesian meta-analyses of observational studies. Starting certainty for all outcomes is **LOW** (observational evidence).

## Summary {#sec-grade-summary}

```{r}
#| label: tbl-grade-summary
#| tbl-cap: "GRADE certainty of evidence — all outcomes, both comparisons."

tibble::tribble(
  ~Outcome, ~Comp, ~k, ~N, ~`OR [95% CrI]`, ~Certainty, ~`Key Limitation`,
  "OS",            "C1", 35, 17661, "0.77 [0.71\u20130.83]", "\u2295\u2295\u2296\u2296 LOW",       "I\u00B2 = 68%; obs only",
  "OS",            "C2", 10, 10624, "0.69 [0.61\u20130.78]", "\u2295\u2295\u2296\u2296 LOW",       "I\u00B2 = 90% (freq); one registry dominant",
  "NRM",           "C1", 15,  2081, "0.86 [0.70\u20131.08]", "\u2295\u2296\u2296\u2296 V.LOW",     "Complete M2 mediation; CrI crosses null",
  "NRM",           "C2",  5,  1186, "0.65 [0.43\u20130.97]", "\u2295\u2296\u2296\u2296 V.LOW",     "k = 5; M2 attenuation; CrI fragile",
  "aGVHD II\u2013IV", "C1", 24,  3837, "0.63 [0.54\u20130.74]", "\u2295\u2295\u2296\u2296 LOW",   "I\u00B2 = 63%; obs only",
  "aGVHD II\u2013IV", "C2",  6,   784, "0.59 [0.42\u20130.84]", "\u2295\u2295\u2296\u2296 LOW",   "k = 6; \u03c4 poorly estimated",
  "CMV",           "C1", 18,  3357, "1.41 [1.20\u20131.65]", "\u2295\u2295\u2296\u2296 LOW",       "I\u00B2 = 69%; obs only",
  "CMV",           "C2", 12,  1267, "0.77 [0.60\u20130.99]", "\u2295\u2296\u2296\u2296 V.LOW",     "CrI fragile; direction reversal is key",
  "BSI",           "C1",  6,  1319, "2.39 [1.85\u20133.10]", "\u2295\u2296\u2296\u2296 V.LOW",     "k = 6; 7 studies NR excluded",
  "BSI",           "C2",  3,   613, "2.03 [1.33\u20133.09]\u2020", "(\u2295\u2296\u2296\u2296)",   "k = 3; definitional heterogeneity",
  "IFI",           "C1",  8,  3816, "0.73 [0.53\u20131.00]", "\u2295\u2296\u2296\u2296 V.LOW",     "\u03c4 = 1.05 (Yanada); CrI \u2248 null",
  "IFI",           "C2", NA,    NA, "\u2014",                 "\u2014",                              "Insufficient data"
) |>
  kable(align = c("l","c","r","r","c","c","l")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) |>
  footnote(symbol = c("Mantel-Haenszel fixed-effect; formal GRADE not performed at k = 3."))
```

## Domain-Level Assessments {#sec-grade-domains}

### Overall Survival

```{r}
#| label: tbl-grade-os
#| tbl-cap: "GRADE domain assessment — Overall Survival."

tibble::tribble(
  ~Domain, ~C1, ~C2,
  "Study design", "34 retro/registry + 1 RCT; start LOW", "All retro/registry; start LOW",
  "Risk of bias", "No downgrade (effect survives M2)", "No downgrade (same reasoning)",
  "Inconsistency", "\u22121 (I\u00B2 = 68%)", "\u22121 (I\u00B2 = 90%, two extreme outliers)",
  "Indirectness", "0", "0",
  "Imprecision", "0 (CrI below null; N = 17,661)", "0 (CrI below null; N = 10,624)",
  "Pub. bias", "0", "0",
  "Upgrade", "0", "+1 (cross-comparison coherence)",
  "Net", "\u22121", "0",
  "CERTAINTY", "\u2295\u2295\u2296\u2296 LOW", "\u2295\u2295\u2296\u2296 LOW"
) |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

### Non-Relapse Mortality

```{r}
#| label: tbl-grade-nrm
#| tbl-cap: "GRADE domain assessment — NRM."

tibble::tribble(
  ~Domain, ~C1, ~C2,
  "Study design", "All retro/registry; start LOW", "All retro/registry; start LOW",
  "Risk of bias", "\u22121 (complete M2 mediation = confounding signal)", "\u22121 (M2 attenuation, CrI crosses null)",
  "Inconsistency", "0 (I\u00B2 = 39%)", "0 (I\u00B2 = 0%)",
  "Indirectness", "0", "0",
  "Imprecision", "\u22121 (CrI crosses null; M2 CrI fully crosses)", "\u22121 (k = 5; M2 CrI crosses null)",
  "Pub. bias", "0", "0",
  "Net", "\u22122", "\u22122",
  "CERTAINTY", "\u2295\u2296\u2296\u2296 V.LOW", "\u2295\u2296\u2296\u2296 V.LOW"
) |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

### aGVHD Grade II–IV

```{r}
#| label: tbl-grade-agvhd
#| tbl-cap: "GRADE domain assessment — aGVHD."

tibble::tribble(
  ~Domain, ~C1, ~C2,
  "Study design", "All retro/registry; start LOW", "All retro/registry; start LOW",
  "Risk of bias", "0 (bias opposes effect)", "0 (same reasoning)",
  "Inconsistency", "\u22121 (I\u00B2 = 63%, \u03c4 = 0.66)", "\u22121 (I\u00B2 = 67%; \u03c4 poorly estimated)",
  "Indirectness", "0", "0",
  "Imprecision", "0 (CrI below null)", "0 (CrI below null)",
  "Upgrade", "+1 (confounders oppose)", "+1 (confounders oppose + cross-comparison)",
  "Net", "0", "0",
  "CERTAINTY", "\u2295\u2295\u2296\u2296 LOW", "\u2295\u2295\u2296\u2296 LOW"
) |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

### CMV Reactivation

```{r}
#| label: tbl-grade-cmv
#| tbl-cap: "GRADE domain assessment — CMV."

tibble::tribble(
  ~Domain, ~C1, ~C2,
  "Study design", "All retrospective; start LOW", "All retrospective; start LOW",
  "Risk of bias", "0 (ascertainment bias noted)", "0 (direction uncertain)",
  "Inconsistency", "\u22121 (I\u00B2 = 69%)", "\u22121 (I\u00B2 = 67%)",
  "Indirectness", "0", "0",
  "Imprecision", "0 (CrI excludes null)", "\u22121 (CrI fragile; crosses null in M2 and sensitivity)",
  "Upgrade", "+1 (steroid-independence confirmed)", "+1 (mechanistic coherence with C1 reversal)",
  "Net", "0", "\u22121",
  "CERTAINTY", "\u2295\u2295\u2296\u2296 LOW", "\u2295\u2296\u2296\u2296 V.LOW"
) |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

::: {.callout-important}
## CMV: The Reversal Is More Informative Than Either Estimate
Neither the C1 CMV estimate (OR 1.41, LOW certainty) nor the C2 estimate (OR 0.77, VERY LOW certainty) alone provides high confidence. However, the *contrast* between them — PTCy harmful vs CNI+MTX/MMF but protective vs ATG — provides **moderate-certainty evidence** that CMV risk is ordered by T-cell depletion depth. This cross-comparison finding has higher certainty than either component.
:::

## Narrative Summary {#sec-grade-narrative}

Across both comparisons, certainty does not exceed LOW — a predictable consequence of an almost entirely observational evidence base.

**Cross-comparison coherence** is the strongest methodological asset. OS and aGVHD show consistent effects across two fundamentally different comparators, reducing the probability that findings are artefacts of comparator-specific confounding. The mediation pattern also replicates (OS: partial; NRM: substantial/complete).

**The CMV direction reversal** provides moderate-certainty mechanistic evidence. Any C2 value below ~1.0 confirms a qualitative direction shift, making this finding robust to the specific C2 point estimate.

**Infection outcomes remain the principal evidence gap.** BSI and IFI carry VERY LOW certainty in both comparisons. Dedicated prospective studies with standardised endpoint definitions are needed.
