---
title: "GRADE Certainty Assessment"
---

```{r}
#| label: setup
#| include: false
source("_common.R")

# Load Table2 for dynamic values
t2 <- read_csv(file.path(model_dir, "Table2_post_block9.csv"), show_col_types = FALSE)
# Set B frequentist concordance results (23 outcomes). Columns differ from the
# superseded post_block9/freq_results.csv: freq_lo/freq_hi/tau_freq/Q_p.
freq <- read_csv(resolve_model_file("freq_results_setB.csv"), show_col_types = FALSE)
```

GRADE certainty was assessed per the GRADE Handbook (Schünemann et al.), adapted for Bayesian meta-analyses of observational studies. Starting certainty for all outcomes is **LOW** (observational evidence). Post-Block-9 corpus: 251 studies (14 RCTs), 525 arms.

## Summary {#sec-grade-summary}

```{r}
#| label: tbl-grade-summary
#| tbl-cap: "GRADE certainty of evidence — all outcomes, both comparisons. Updated for post-Block-9 results (May 2026)."

tibble::tribble(
  ~Outcome, ~Comp, ~k, ~N, ~`OR [95% CrI]`, ~Certainty, ~`Key Limitation`,
  "OS",            "C1", 35, 10133, "0.84 [0.76\u20130.92]", "\u2295\u2295\u2296\u2296 LOW",       "I\u00B2 = 64%; obs only",
  "OS",            "C2",  9, 12791, "0.83 [0.75\u20130.92]", "\u2295\u2295\u2296\u2296 LOW",       "I\u00B2 = 93% (freq); attenuated from pre-B9",
  "NRM",           "C1", 12,  1309, "0.88 [0.66\u20131.18]", "\u2295\u2296\u2296\u2296 V.LOW",     "Complete M2 mediation; CrI crosses null",
  "NRM",           "C2", NA,    NA, "\u2014 (k = 2)",         "\u2014",                              "Below minimum k for modelling",
  "aGVHD II\u2013IV", "C1", 28,  4197, "0.67 [0.59\u20130.78]", "\u2295\u2296\u2296\u2296 V.LOW", "Egger p = 0.04; T&F adj OR 0.83",
  "aGVHD II\u2013IV", "C2",  8,  1039, "0.63 [0.47\u20130.83]", "\u2295\u2295\u2296\u2296 LOW",   "I\u00B2 = 73%; k = 9",
  "CMV",           "C1", 22,  3328, "1.26 [1.07\u20131.47]", "\u2295\u2295\u2296\u2296 LOW",       "I\u00B2 = 62%; obs only",
  "CMV",           "C2", 12,  1511, "0.97 [0.77\u20131.23]", "\u2295\u2296\u2296\u2296 V.LOW",     "CrI crosses null; sensitivity convergence",
  "BSI",           "C1",  6,   806, "1.87 [1.33\u20132.62]", "\u2295\u2296\u2296\u2296 V.LOW",     "k = 6; freq CrI crosses null",
  "IFI",           "C1",  6,  2243, "0.43 [0.29\u20130.63]", "\u2295\u2296\u2296\u2296 V.LOW",     "\u03c4 = 1.53; extreme heterogeneity",
  "cGVHD mod-sev", "C1", 19,  2840, "0.47 [0.38\u20130.57]", "\u2295\u2295\u2296\u2296 LOW",       "Largest effect; I\u00B2 = 80%",
  "cGVHD mod-sev", "C2",  6,   905, "0.79 [0.55\u20131.14]", "\u2295\u2296\u2296\u2296 V.LOW",     "CrI crosses null; k = 6",
  "RRM",           "C1", 34, 11121, "0.86 [0.77\u20130.97]", "\u2295\u2295\u2296\u2296 LOW",       "Consistent; no steroid mediation",
  "RRM",           "C2", 10, 17182, "0.79 [0.69\u20130.89]", "\u2295\u2295\u2296\u2296 LOW",       "\u03c4 = 0.14; highly consistent",
  "BK Virus",      "C1", 11,  1333, "2.48 [1.82\u20133.38]", "\u2295\u2296\u2296\u2296 V.LOW",     "\u03c4 = 1.00; dual mechanism",
  "BK Virus",      "C2",  3,   313, "2.44 [1.40\u20134.26]", "\u2295\u2296\u2296\u2296 V.LOW",     "k = 3; hypothesis-generating",
  "IRM",           "C1", 13,  5504, "1.19 [1.00\u20131.43]", "\u2295\u2296\u2296\u2296 V.LOW",     "Bayes-freq divergence; CIBMTR-driven",
  "IRM",           "C2",  7, 13346, "0.90 [0.74\u20131.09]", "\u2295\u2296\u2296\u2296 V.LOW",     "CrI crosses null; k = 7"
) |>
  kable(align = c("l","c","r","r","c","c","l")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

::: {.callout-note appearance="simple"}
**Changes from pre-Block-9.** C2 NRM can no longer be graded (k = 2 after arm classification corrections). C2 CMV moved from borderline significant (OR 0.77 [0.60–0.99]) to null-crossing (OR 0.97 [0.77–1.23]). C2 BSI removed (k = 3 Mantel-Haenszel estimate was not formally gradeable). All C1 outcomes stable in direction and certainty.
:::

## Domain-Level Assessments {#sec-grade-domains}

### Overall Survival

```{r}
#| label: tbl-grade-os
#| tbl-cap: "GRADE domain assessment — Overall Survival."

tibble::tribble(
  ~Domain, ~C1, ~C2,
  "Study design", "39 retro/registry + 1 RCT; start LOW", "All retro/registry; start LOW",
  "Risk of bias", "No downgrade (effect survives M2)", "No downgrade (M2 OR 0.82 still protective)",
  "Inconsistency", "\u22121 (I\u00B2 = 64%; \u03c4 = 0.61)", "\u22121 (I\u00B2 = 93%; one registry dominant)",
  "Indirectness", "0", "0",
  "Imprecision", "0 (CrI excludes null; N = 19,724)", "0 (CrI excludes null; N = 12,851)",
  "Pub. bias", "0", "0",
  "Upgrade", "0", "+1 (cross-comparison coherence; sensitivity k=21 strengthens)",
  "Net", "\u22121", "0",
  "CERTAINTY", "\u2295\u2295\u2296\u2296 LOW", "\u2295\u2295\u2296\u2296 LOW"
) |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

The C2 OS sensitivity (expanded arm inclusion, k = 21) strengthened the effect from OR 0.83 to 0.76 [0.68–0.85], supporting the upgrade for cross-comparison coherence.

### Non-Relapse Mortality

```{r}
#| label: tbl-grade-nrm
#| tbl-cap: "GRADE domain assessment — NRM."

tibble::tribble(
  ~Domain, ~C1, ~C2,
  "Study design", "All retro/registry; start LOW", "\u2014",
  "Risk of bias", "\u22121 (M2 shows steroid mediation \u2192 confounding signal)", "\u2014",
  "Inconsistency", "0 (I\u00B2 = 0%)", "\u2014",
  "Indirectness", "0", "\u2014",
  "Imprecision", "\u22121 (CrI crosses null; k = 12)", "\u2014",
  "Pub. bias", "0", "\u2014",
  "Net", "\u22122", "\u2014",
  "CERTAINTY", "\u2295\u2296\u2296\u2296 V.LOW", "\u2014 (k = 2; not gradeable)"
) |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

C2 NRM dropped below the modelling threshold (k = 2) after removing Sayin 2024 (misclassified comparator). The pre-Block-9 C2 NRM estimate (k = 5, OR 0.65 [0.43–0.97]) should be interpreted with caution given the arm classification issues that reduced k.

### aGVHD Grade II–IV

```{r}
#| label: tbl-grade-agvhd
#| tbl-cap: "GRADE domain assessment — aGVHD."

tibble::tribble(
  ~Domain, ~C1, ~C2,
  "Study design", "All retro/registry; start LOW", "All retro/registry; start LOW",
  "Risk of bias", "0 (ascertainment bias opposes effect)", "0 (same reasoning)",
  "Inconsistency", "\u22121 (I\u00B2 = 62%; \u03c4 = 0.67)", "\u22121 (I\u00B2 = 73%; \u03c4 = 0.53)",
  "Indirectness", "0", "0",
  "Imprecision", "0 (CrI excludes null; k = 28)", "0 (CrI excludes null; k = 9)",
  "Pub. bias", "\u22121 (Egger p = 0.041; T&F: 0.61 \u2192 0.83)", "0 (not assessable, k = 9)",
  "Upgrade", "+1 (confounders oppose effect; cross-comparison)", "+1 (confounders oppose; cross-comparison; sens k=14 consistent)",
  "Net", "\u22121", "0",
  "CERTAINTY", "\u2295\u2296\u2296\u2296 V.LOW", "\u2295\u2295\u2296\u2296 LOW"
) |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

C1 aGVHD shows significant funnel plot asymmetry (Egger's p = 0.041), with small studies disproportionately favouring PTCy. Trim-and-fill imputes 8 studies, shifting the frequentist OR from 0.61 to 0.83 [0.61–1.13]. The Bayesian estimate (OR 0.67 [0.59–0.78]) is less susceptible to small-study effects due to shrinkage but does not fully address selective publication. C1 aGVHD is accordingly downgraded to **VERY LOW**. The C2 aGVHD estimate (OR 0.63, k = 8) is too small for formal bias assessment but remains consistent.

### CMV Reactivation

```{r}
#| label: tbl-grade-cmv
#| tbl-cap: "GRADE domain assessment — CMV."

tibble::tribble(
  ~Domain, ~C1, ~C2,
  "Study design", "1 RCT + 21 retro; start LOW", "All retrospective; start LOW",
  "Risk of bias", "0 (monitoring protocol heterogeneity noted)", "0 (direction uncertain)",
  "Inconsistency", "\u22121 (I\u00B2 = 62%)", "\u22121 (I\u00B2 = 71%)",
  "Indirectness", "0", "0",
  "Imprecision", "0 (CrI excludes null [1.07\u20131.47])", "\u22121 (CrI crosses null [0.73\u20131.15])",
  "Upgrade", "+1 (steroid-independence; sensitivity convergence pattern)", "0 (see below)",
  "Net", "0", "\u22122",
  "CERTAINTY", "\u2295\u2295\u2296\u2296 LOW", "\u2295\u2296\u2296\u2296 V.LOW"
) |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

::: {.callout-important}
## CMV: From Direction Reversal to Depletion Gradient

The pre-Block-9 analysis reported a qualitative direction reversal: PTCy increased CMV vs CNI+MTX (OR 1.41) but decreased it vs ATG (OR 0.77 [0.60–0.99]). In the post-Block-9 corpus, C2 CMV has shifted to OR 0.97 [0.77–1.23], and the CrI crosses null. The binary reversal is no longer statistically credible.

However, the **C2 sensitivity analysis** provides a different — and arguably stronger — form of evidence. When PTCy+ATG combination arms are reclassified as "PTCy-side" (k = 22), the CMV estimate converges to OR 1.00 [0.85–1.18]. This convergence is predicted by the depletion hierarchy: when both arms have deep T-cell depletion, CMV rates should equalise.

The three estimates trace a gradient:

| Analysis | Arms compared | OR | Interpretation |
|---|---|---|---|
| C1 primary | PTCy vs CNI | 1.26 | PTCy depletes more → more CMV |
| C2 primary | PTCy vs ATG | 0.97 | Near-parity; slight PTCy advantage |
| C2 sensitivity | PTCy+ATG vs ATG | 1.00 | Equal depletion → equal CMV |

This gradient is more consistent with the depletion hierarchy than the pre-Block-9 binary reversal. The mechanistic conclusion — that CMV risk tracks T-cell depletion depth, not any PTCy-specific property — carries **LOW certainty** when assessed across the full evidence pattern, despite the individual C2 estimate being VERY LOW.
:::

### BSI and IFI

```{r}
#| label: tbl-grade-infection
#| tbl-cap: "GRADE domain assessment — BSI and IFI."

tibble::tribble(
  ~Domain, ~`C1 BSI`, ~`C1 IFI`,
  "Study design", "All retro/registry; start LOW", "All retro/registry; start LOW",
  "Risk of bias", "\u22121 (definitions heterogeneous)", "0",
  "Inconsistency", "\u22121 (I\u00B2 = 71%; freq CrI crosses null)", "\u22121 (\u03c4 = 1.53; I\u00B2 = 80%)",
  "Indirectness", "0", "0",
  "Imprecision", "0 (Bayes CrI excludes null)", "0 (CrI excludes null [0.29\u20130.63])",
  "Net", "\u22122", "\u22121",
  "CERTAINTY", "\u2295\u2296\u2296\u2296 V.LOW", "\u2295\u2296\u2296\u2296 V.LOW"
) |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

BSI: The Bayesian OR of 1.87 [1.33–2.62] suggests elevated risk, but the frequentist estimate (OR 1.29 [0.61–2.73]) crosses null substantially. Both k = 6. The Bayesian shrinkage prior and the frequentist random-effects model disagree, reducing confidence.

IFI: The post-Block-9 estimate (OR 0.43 [0.29–0.63]) is substantially stronger than pre-Block-9 (0.73 [0.53–1.00]), but τ = 1.53 indicates extreme between-study heterogeneity. The frequentist CrI (0.21–1.74) again crosses null. With only k = 6 studies and this level of heterogeneity, the strong protective signal should be interpreted cautiously.

::: {.callout-warning}
## Publication-bias analyses are pre-deduplication

The Egger, trim-and-fill and RoBMA results in this section were computed before the
Set B cohort deduplication and have not been regenerated. They should be read as
indicative only; the C1 OS funnel in particular included five EBMT ALWP publications
that are now represented by one.
:::

## Publication Bias {#sec-grade-pubbias}

Formal assessment was conducted for C1 outcomes with k ≥ 10 using Egger's regression test and trim-and-fill.

```{r}
#| label: tbl-grade-pubbias
#| tbl-cap: "Publication bias assessment — C1 outcomes with k ≥ 10."

tibble::tribble(
  ~Outcome, ~k, ~`Egger z`, ~`Egger p`, ~`T&F imputed`, ~`Original OR`, ~`Adjusted OR`,
  "OS",      40, -1.02, 0.314, 7, "0.77 [0.66\u20130.90]", "0.86 [0.73\u20131.02]",
  "aGVHD",   28, -2.15, 0.041, 8, "0.61 [0.47\u20130.79]", "0.83 [0.61\u20131.13]",
  "CMV",     22, -0.21, 0.835, 0, "1.25 [0.94\u20131.66]", "1.25 [0.94\u20131.66]"
) |>
  mutate(`Egger p` = ifelse(`Egger p` < 0.05,
                             paste0("**", sprintf("%.3f", `Egger p`), "**"),
                             sprintf("%.3f", `Egger p`))) |>
  kable(align = c("l","r","r","r","r","c","c")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

::: {.callout-warning}
## C1 aGVHD: Significant Funnel Asymmetry
Egger's test detects significant small-study bias (p = 0.041). After trim-and-fill imputation, the adjusted OR shifts from 0.61 to 0.83 and the CI crosses null. The Bayesian pooled estimate (OR 0.67 [0.59–0.78]) is less extreme than the frequentist 0.61, reflecting shrinkage, but does not account for selective non-publication. C1 aGVHD certainty is downgraded from LOW to VERY LOW.

C1 OS shows borderline concern: 7 studies imputed with adjusted OR touching null (0.86 [0.73–1.02]), though Egger's test is non-significant (p = 0.31). C1 CMV is symmetric with no imputed studies.
:::

## Narrative Summary {#sec-grade-narrative}

Across both comparisons, certainty does not exceed LOW — a predictable consequence of an almost entirely observational evidence base.

**The most robust findings** are C1 OS (OR 0.84, k = 35, LOW certainty) and C1 moderate-severe cGVHD (OR 0.47, k = 19, LOW certainty). Both are stable, have CrIs excluding null, and are consistent across comparators. C1 cGVHD is the largest treatment effect in the analysis. C1 RRM (OR 0.86, k = 34, LOW certainty) is similarly robust and steroid-independent, supporting preserved GVL activity. **Cross-comparison coherence** remains the strongest methodological asset.

**Chronic GVHD is arguably the most clinically important finding.** The OR of 0.47 for moderate-severe cGVHD translates to roughly a halving of the leading cause of late post-transplant morbidity. Despite I² = 80%, the Bayesian and frequentist estimates agree on direction and magnitude, and the result is consistent across the any-NIH (OR 0.23) and moderate-severe (OR 0.47) subsets. The previously reported OR of 0.33 came from a dataset built outside the audited pipeline that included two studies ineligible under the cohort-selection rule; see [Results C1](results-c1.qmd).

**The infection–mortality disconnect is the key risk-benefit finding.** PTCy increases infection rates (CMV, BSI, BK) but infection-related mortality is not clearly increased. The C1 IRM Bayesian–frequentist divergence (OR 1.19 vs 1.03) warrants caution, but even the more pessimistic estimate represents a far smaller effect than the infection rate increases, suggesting infectious complications are clinically manageable.

**C2 NRM is no longer gradeable.** Arm classification corrections reduced k to 3. NRM steroid mediation is inconclusive in the post-Block-9 data.

::: {.callout-tip}
## Post-Block-9 GRADE Changes
- **New outcomes added:** cGVHD moderate-severe (C1 LOW, C2 V.LOW), RRM (C1 LOW, C2 LOW), BK virus (V.LOW), IRM (V.LOW)
- C2 NRM: reduced to k = 3 (from k = 5); not formally gradeable
- C2 CMV: still VERY LOW; revised from "direction reversal" to "gradient pattern"
- C1 IFI: k decreased (8 → 6); τ = 1.53 warrants extreme caution
- C1 IRM: Bayesian–frequentist divergence; rated VERY LOW
- C1 NRM mediation: no longer shows clear complete mediation (M1→M2 shift minimal)
:::
