---
title: "GRADE Certainty Assessment"
---

```{r}
#| label: setup
#| include: false
source("_common.R")

# Load Table2 for dynamic values
t2 <- read_csv(file.path(model_dir, "Table2_post_block9.csv"), show_col_types = FALSE)
freq <- read_csv(file.path(model_dir, "freq_results.csv"), show_col_types = FALSE)
```

GRADE certainty was assessed per the GRADE Handbook (Schünemann et al.), adapted for Bayesian meta-analyses of observational studies. Starting certainty for all outcomes is **LOW** (observational evidence). Post-Block-9 corpus: 251 studies (14 RCTs), 525 arms.

## Summary {#sec-grade-summary}

```{r}
#| label: tbl-grade-summary
#| tbl-cap: "GRADE certainty of evidence — all outcomes, both comparisons. Updated for post-Block-9 results (May 2026)."

tibble::tribble(
  ~Outcome, ~Comp, ~k, ~N, ~`OR [95% CrI]`, ~Certainty, ~`Key Limitation`,
  "OS",            "C1", 40, 19724, "0.79 [0.73\u20130.85]", "\u2295\u2295\u2296\u2296 LOW",       "I\u00B2 = 64%; obs only",
  "OS",            "C2", 10, 12851, "0.81 [0.74\u20130.90]", "\u2295\u2295\u2296\u2296 LOW",       "I\u00B2 = 95% (freq); attenuated from pre-B9",
  "NRM",           "C1", 12,  1309, "0.88 [0.66\u20131.18]", "\u2295\u2296\u2296\u2296 V.LOW",     "Complete M2 mediation; CrI crosses null",
  "NRM",           "C2", NA,    NA, "\u2014 (k = 2)",         "\u2014",                              "Below minimum k for modelling",
  "aGVHD II\u2013IV", "C1", 28,  4197, "0.67 [0.59\u20130.78]", "\u2295\u2296\u2296\u2296 V.LOW", "Egger p = 0.04; T&F adj OR 0.83",
  "aGVHD II\u2013IV", "C2",  9,  1099, "0.58 [0.44\u20130.77]", "\u2295\u2295\u2296\u2296 LOW",   "I\u00B2 = 73%; k = 9",
  "CMV",           "C1", 22,  3328, "1.26 [1.07\u20131.47]", "\u2295\u2295\u2296\u2296 LOW",       "I\u00B2 = 62%; obs only",
  "CMV",           "C2", 13,  1571, "0.92 [0.73\u20131.15]", "\u2295\u2296\u2296\u2296 V.LOW",     "CrI crosses null; sensitivity convergence",
  "BSI",           "C1",  6,   806, "1.87 [1.33\u20132.62]", "\u2295\u2296\u2296\u2296 V.LOW",     "k = 6; freq CrI crosses null",
  "IFI",           "C1",  6,  2243, "0.43 [0.29\u20130.63]", "\u2295\u2296\u2296\u2296 V.LOW",     "\u03c4 = 1.64; extreme heterogeneity"
) |>
  kable(align = c("l","c","r","r","c","c","l")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

::: {.callout-note appearance="simple"}
**Changes from pre-Block-9.** C2 NRM can no longer be graded (k = 2 after arm classification corrections). C2 CMV moved from borderline significant (OR 0.77 [0.60–0.99]) to null-crossing (OR 0.92 [0.73–1.15]). C2 BSI removed (k = 3 Mantel-Haenszel estimate was not formally gradeable). All C1 outcomes stable in direction and certainty.
:::

## Domain-Level Assessments {#sec-grade-domains}

### Overall Survival

```{r}
#| label: tbl-grade-os
#| tbl-cap: "GRADE domain assessment — Overall Survival."

tibble::tribble(
  ~Domain, ~C1, ~C2,
  "Study design", "39 retro/registry + 1 RCT; start LOW", "All retro/registry; start LOW",
  "Risk of bias", "No downgrade (effect survives M2)", "No downgrade (M2 OR 0.82 still protective)",
  "Inconsistency", "\u22121 (I\u00B2 = 64%; \u03c4 = 0.61)", "\u22121 (I\u00B2 = 95%; one registry dominant)",
  "Indirectness", "0", "0",
  "Imprecision", "0 (CrI excludes null; N = 19,724)", "0 (CrI excludes null; N = 12,851)",
  "Pub. bias", "0", "0",
  "Upgrade", "0", "+1 (cross-comparison coherence; sensitivity k=21 strengthens)",
  "Net", "\u22121", "0",
  "CERTAINTY", "\u2295\u2295\u2296\u2296 LOW", "\u2295\u2295\u2296\u2296 LOW"
) |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

The C2 OS sensitivity (expanded arm inclusion, k = 21) strengthened the effect from OR 0.81 to 0.74 [0.66–0.83], supporting the upgrade for cross-comparison coherence.

### Non-Relapse Mortality

```{r}
#| label: tbl-grade-nrm
#| tbl-cap: "GRADE domain assessment — NRM."

tibble::tribble(
  ~Domain, ~C1, ~C2,
  "Study design", "All retro/registry; start LOW", "\u2014",
  "Risk of bias", "\u22121 (M2 shows steroid mediation \u2192 confounding signal)", "\u2014",
  "Inconsistency", "0 (I\u00B2 = 0%)", "\u2014",
  "Indirectness", "0", "\u2014",
  "Imprecision", "\u22121 (CrI crosses null; k = 12)", "\u2014",
  "Pub. bias", "0", "\u2014",
  "Net", "\u22122", "\u2014",
  "CERTAINTY", "\u2295\u2296\u2296\u2296 V.LOW", "\u2014 (k = 2; not gradeable)"
) |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

C2 NRM dropped below the modelling threshold (k = 2) after removing Sayin 2024 (misclassified comparator). The pre-Block-9 C2 NRM estimate (k = 5, OR 0.65 [0.43–0.97]) should be interpreted with caution given the arm classification issues that reduced k.

### aGVHD Grade II–IV

```{r}
#| label: tbl-grade-agvhd
#| tbl-cap: "GRADE domain assessment — aGVHD."

tibble::tribble(
  ~Domain, ~C1, ~C2,
  "Study design", "All retro/registry; start LOW", "All retro/registry; start LOW",
  "Risk of bias", "0 (ascertainment bias opposes effect)", "0 (same reasoning)",
  "Inconsistency", "\u22121 (I\u00B2 = 62%; \u03c4 = 0.68)", "\u22121 (I\u00B2 = 73%; \u03c4 = 0.55)",
  "Indirectness", "0", "0",
  "Imprecision", "0 (CrI excludes null; k = 28)", "0 (CrI excludes null; k = 9)",
  "Pub. bias", "\u22121 (Egger p = 0.041; T&F: 0.61 \u2192 0.83)", "0 (not assessable, k = 9)",
  "Upgrade", "+1 (confounders oppose effect; cross-comparison)", "+1 (confounders oppose; cross-comparison; sens k=14 consistent)",
  "Net", "\u22121", "0",
  "CERTAINTY", "\u2295\u2296\u2296\u2296 V.LOW", "\u2295\u2295\u2296\u2296 LOW"
) |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

C1 aGVHD shows significant funnel plot asymmetry (Egger's p = 0.041), with small studies disproportionately favouring PTCy. Trim-and-fill imputes 8 studies, shifting the frequentist OR from 0.61 to 0.83 [0.61–1.13]. The Bayesian estimate (OR 0.67 [0.59–0.78]) is less susceptible to small-study effects due to shrinkage but does not fully address selective publication. C1 aGVHD is accordingly downgraded to **VERY LOW**. The C2 aGVHD estimate (OR 0.58, k = 9) is too small for formal bias assessment but remains consistent.

### CMV Reactivation

```{r}
#| label: tbl-grade-cmv
#| tbl-cap: "GRADE domain assessment — CMV."

tibble::tribble(
  ~Domain, ~C1, ~C2,
  "Study design", "1 RCT + 21 retro; start LOW", "All retrospective; start LOW",
  "Risk of bias", "0 (monitoring protocol heterogeneity noted)", "0 (direction uncertain)",
  "Inconsistency", "\u22121 (I\u00B2 = 62%)", "\u22121 (I\u00B2 = 71%)",
  "Indirectness", "0", "0",
  "Imprecision", "0 (CrI excludes null [1.07\u20131.47])", "\u22121 (CrI crosses null [0.73\u20131.15])",
  "Upgrade", "+1 (steroid-independence; sensitivity convergence pattern)", "0 (see below)",
  "Net", "0", "\u22122",
  "CERTAINTY", "\u2295\u2295\u2296\u2296 LOW", "\u2295\u2296\u2296\u2296 V.LOW"
) |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

::: {.callout-important}
## CMV: From Direction Reversal to Depletion Gradient

The pre-Block-9 analysis reported a qualitative direction reversal: PTCy increased CMV vs CNI+MTX (OR 1.41) but decreased it vs ATG (OR 0.77 [0.60–0.99]). In the post-Block-9 corpus, C2 CMV has shifted to OR 0.92 [0.73–1.15], and the CrI crosses null. The binary reversal is no longer statistically credible.

However, the **C2 sensitivity analysis** provides a different — and arguably stronger — form of evidence. When PTCy+ATG combination arms are reclassified as "PTCy-side" (k = 22), the CMV estimate converges to OR 1.00 [0.85–1.18]. This convergence is predicted by the depletion hierarchy: when both arms have deep T-cell depletion, CMV rates should equalise.

The three estimates trace a gradient:

| Analysis | Arms compared | OR | Interpretation |
|---|---|---|---|
| C1 primary | PTCy vs CNI | 1.26 | PTCy depletes more → more CMV |
| C2 primary | PTCy vs ATG | 0.92 | Near-parity; slight PTCy advantage |
| C2 sensitivity | PTCy+ATG vs ATG | 1.00 | Equal depletion → equal CMV |

This gradient is more consistent with the depletion hierarchy than the pre-Block-9 binary reversal. The mechanistic conclusion — that CMV risk tracks T-cell depletion depth, not any PTCy-specific property — carries **LOW certainty** when assessed across the full evidence pattern, despite the individual C2 estimate being VERY LOW.
:::

### BSI and IFI

```{r}
#| label: tbl-grade-infection
#| tbl-cap: "GRADE domain assessment — BSI and IFI."

tibble::tribble(
  ~Domain, ~`C1 BSI`, ~`C1 IFI`,
  "Study design", "All retro/registry; start LOW", "All retro/registry; start LOW",
  "Risk of bias", "\u22121 (definitions heterogeneous)", "0",
  "Inconsistency", "\u22121 (I\u00B2 = 71%; freq CrI crosses null)", "\u22121 (\u03c4 = 1.64; I\u00B2 = 80%)",
  "Indirectness", "0", "0",
  "Imprecision", "0 (Bayes CrI excludes null)", "0 (CrI excludes null [0.29\u20130.63])",
  "Net", "\u22122", "\u22121",
  "CERTAINTY", "\u2295\u2296\u2296\u2296 V.LOW", "\u2295\u2296\u2296\u2296 V.LOW"
) |>
  kable() |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

BSI: The Bayesian OR of 1.87 [1.33–2.62] suggests elevated risk, but the frequentist estimate (OR 1.29 [0.61–2.73]) crosses null substantially. Both k = 6. The Bayesian shrinkage prior and the frequentist random-effects model disagree, reducing confidence.

IFI: The post-Block-9 estimate (OR 0.43 [0.29–0.63]) is substantially stronger than pre-Block-9 (0.73 [0.53–1.00]), but τ = 1.64 indicates extreme between-study heterogeneity. The frequentist CrI (0.21–1.74) again crosses null. With only k = 6 studies and this level of heterogeneity, the strong protective signal should be interpreted cautiously.

## Publication Bias {#sec-grade-pubbias}

Formal assessment was conducted for C1 outcomes with k ≥ 10 using Egger's regression test and trim-and-fill.

```{r}
#| label: tbl-grade-pubbias
#| tbl-cap: "Publication bias assessment — C1 outcomes with k ≥ 10."

tibble::tribble(
  ~Outcome, ~k, ~`Egger z`, ~`Egger p`, ~`T&F imputed`, ~`Original OR`, ~`Adjusted OR`,
  "OS",      40, -1.02, 0.314, 7, "0.77 [0.66\u20130.90]", "0.86 [0.73\u20131.02]",
  "aGVHD",   28, -2.15, 0.041, 8, "0.61 [0.47\u20130.79]", "0.83 [0.61\u20131.13]",
  "CMV",     22, -0.21, 0.835, 0, "1.25 [0.94\u20131.66]", "1.25 [0.94\u20131.66]"
) |>
  mutate(`Egger p` = ifelse(`Egger p` < 0.05,
                             paste0("**", sprintf("%.3f", `Egger p`), "**"),
                             sprintf("%.3f", `Egger p`))) |>
  kable(align = c("l","r","r","r","r","c","c")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

::: {.callout-warning}
## C1 aGVHD: Significant Funnel Asymmetry
Egger's test detects significant small-study bias (p = 0.041). After trim-and-fill imputation, the adjusted OR shifts from 0.61 to 0.83 and the CI crosses null. The Bayesian pooled estimate (OR 0.67 [0.59–0.78]) is less extreme than the frequentist 0.61, reflecting shrinkage, but does not account for selective non-publication. C1 aGVHD certainty is downgraded from LOW to VERY LOW.

C1 OS shows borderline concern: 7 studies imputed with adjusted OR touching null (0.86 [0.73–1.02]), though Egger's test is non-significant (p = 0.31). C1 CMV is symmetric with no imputed studies.
:::

## Narrative Summary {#sec-grade-narrative}

Across both comparisons, certainty does not exceed LOW — a predictable consequence of an almost entirely observational evidence base.

**The most robust finding** is C1 OS (OR 0.79, k = 40, LOW certainty), which has been stable across the pre-Block-9 and post-Block-9 analyses, is consistent across comparators (C2 OR 0.81), and has CrIs that exclude null. C1 aGVHD (OR 0.67, k = 28) is consistent across comparators (C2 OR 0.58, k = 9) and directionally stable, but is downgraded to VERY LOW due to significant funnel asymmetry (Egger p = 0.041). The trim-and-fill adjusted estimate (OR 0.83) crosses null, suggesting the effect size may be inflated by selective reporting of small positive studies. **Cross-comparison coherence** remains the strongest methodological asset.

**The CMV depletion hierarchy** is supported not by a single direction reversal but by a three-point gradient across the primary and sensitivity analyses. This gradient pattern carries greater evidential weight than any one estimate. Certainty for the mechanistic conclusion (that CMV risk tracks depletion depth) is LOW when integrating the full pattern.

**Infection outcomes remain the principal evidence gap.** BSI and IFI carry VERY LOW certainty. Both show substantial discordance between Bayesian and frequentist estimators, and both have extreme heterogeneity. Dedicated prospective studies with standardised endpoint definitions are needed.

**C2 NRM is no longer gradeable.** Arm classification corrections reduced k to 2. The steroid mediation pattern observed in C1 NRM (OR 0.88 → M2 OR 0.90) remains the best available evidence for the GVHD→steroid→NRM causal chain.

::: {.callout-tip}
## Post-Block-9 GRADE Changes
- C2 NRM: removed (k = 2 after arm reclassification)
- C2 BSI: removed (k = 3 Mantel-Haenszel was not formally gradeable)
- C2 CMV: still VERY LOW; reasoning revised from "direction reversal" to "gradient pattern"
- C1 IFI: k decreased (8 → 6), but CrI now excludes null; τ = 1.64 warrants extreme caution
- All other outcomes: certainty unchanged
:::
