---
title: "Post-Transplant Cyclophosphamide Meta-Analysis"
subtitle: "Bayesian random-effects meta-analysis of PTCy vs alternative GVHD prophylaxis platforms"
date: 2026-05-28
---

```{r}
#| label: setup
#| include: false
source("_common.R")

# Load all posterior draws
post_os      <- load_post("post_os.rds")
post_m2_os   <- load_post("post_m2_os.rds")
post_nrm     <- load_post("post_nrm.rds")
post_m2_nrm  <- load_post("post_m2_nrm.rds")
post_agvhd   <- load_post("post_agvhd.rds")
post_m1      <- load_post("post_m1.rds")
post_m2      <- load_post("post_m2.rds")
post_bsi     <- load_post("post_bsi.rds")
post_ifi     <- load_post("post_ifi.rds")

post_c2_os     <- load_post("post_c2_os.rds")
post_c2_m2_os  <- load_post("post_c2_m2_os.rds")
post_c2_nrm    <- load_post("post_c2_nrm.rds")
post_c2_agvhd  <- load_post("post_c2_agvhd.rds")
post_c2_cmv    <- load_post("post_c2_cmv.rds")
post_c2_m2_cmv <- load_post("post_c2_m2_cmv.rds")
```

## Abstract

::: {.callout-note appearance="minimal"}

**Background.** Post-transplant cyclophosphamide (PTCy) has become widely adopted for GVHD prophylaxis in allogeneic haematopoietic cell transplantation, yet comprehensive quantitative evidence synthesis across survival, GVHD, and infection outcomes — and across comparator regimens — has been lacking.

**Methods.** We conducted a Bayesian random-effects meta-analysis of observational studies comparing PTCy to (1) CNI+MTX/MMF-based regimens (Comparison 1; k = 6–35 across outcomes) and (2) ATG-based regimens (Comparison 2; k = 5–12). Primary outcomes included overall survival (OS), non-relapse mortality (NRM), acute GVHD grade II–IV, CMV reactivation, bloodstream infection (BSI), and invasive fungal infection (IFI). Steroid-mediation models (M2) tested whether GVHD-driven steroid exposure confounded observed effects.

**Results.** PTCy was associated with improved OS (C1: OR 0.77 [0.71–0.83]; C2: OR 0.69 [0.61–0.78]) and reduced aGVHD (C1: OR 0.63; C2: OR 0.59) across both comparators. NRM benefit was substantially or completely mediated by steroid exposure in both comparisons. CMV reactivation showed a direction reversal: increased with PTCy vs CNI+MTX/MMF (OR 1.41 [1.20–1.65]) but decreased vs ATG (OR 0.77 [0.60–0.99]), consistent with a T-cell depletion depth hierarchy. BSI risk was elevated in both comparisons (C1: OR 2.39; C2: MH OR 2.03).

**Conclusions.** PTCy provides consistent survival and GVHD advantages regardless of comparator. The CMV direction reversal establishes that infection risk is determined by the relative depth of T-cell depletion, not by any intrinsic PTCy property. NRM benefit is steroid-mediated. GRADE certainty is LOW for OS and aGVHD, VERY LOW for other outcomes.
:::

---

## Executive Summary

```{r}
#| label: fig-exec-summary
#| fig-cap: "Bayesian meta-analysis — PTCy effect across outcomes and comparators. Filled circles represent direct-effect models (M1); open circles represent steroid-adjusted models (M2). The dashed vertical line marks OR = 1 (no effect). OR < 1 favours PTCy for all outcomes."
#| fig-height: 7
#| fig-width: 10

pd <- do.call(rbind, list(
  or_summary(post_os,    "b_ptcy_binary", "OS", "C1: PTCy vs CNI+MTX"),
  or_summary(post_m2_os, "b_eventsn_ptcy_binary", "OS", "C1: PTCy vs CNI+MTX", "M2"),
  or_summary(post_nrm,    "b_ptcy_binary", "NRM", "C1: PTCy vs CNI+MTX"),
  or_summary(post_m2_nrm, "b_eventsn_ptcy_binary", "NRM", "C1: PTCy vs CNI+MTX", "M2"),
  or_summary(post_agvhd, "b_ptcy_binary", "aGVHD II\u2013IV", "C1: PTCy vs CNI+MTX"),
  or_summary(post_m1,    "b_ptcy_binary", "CMV", "C1: PTCy vs CNI+MTX"),
  or_summary(post_m2,    "b_eventsn_ptcy_binary", "CMV", "C1: PTCy vs CNI+MTX", "M2"),
  or_summary(post_bsi,   "b_ptcy_binary", "BSI", "C1: PTCy vs CNI+MTX"),
  or_summary(post_ifi,   "b_ptcy_binary", "IFI", "C1: PTCy vs CNI+MTX"),
  or_summary(post_c2_os,     "b_ptcy_binary", "OS", "C2: PTCy vs ATG"),
  or_summary(post_c2_m2_os,  "b_ptcy_binary", "OS", "C2: PTCy vs ATG", "M2"),
  or_summary(post_c2_nrm,    "b_ptcy_binary", "NRM", "C2: PTCy vs ATG"),
  or_summary(post_c2_agvhd,  "b_ptcy_binary", "aGVHD II\u2013IV", "C2: PTCy vs ATG"),
  or_summary(post_c2_cmv,    "b_ptcy_binary", "CMV", "C2: PTCy vs ATG"),
  or_summary(post_c2_m2_cmv, "b_ptcy_binary", "CMV", "C2: PTCy vs ATG", "M2")
)) |>
  as_tibble() |>
  mutate(
    outcome = factor(outcome, levels = rev(c("OS", "NRM", "aGVHD II\u2013IV", "CMV", "BSI", "IFI"))),
    comparison = factor(comparison, levels = c("C1: PTCy vs CNI+MTX", "C2: PTCy vs ATG")),
    label = sprintf("%.2f [%.2f, %.2f]", or, lo, hi),
    y_num = as.numeric(outcome),
    y_pos = y_num + ifelse(model == "M2", -0.15, 0.15)
  )

ggplot(pd, aes(x = or, y = y_pos, shape = model, fill = model)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = jama_grey, linewidth = 0.4) +
  geom_pointrange(aes(xmin = lo, xmax = hi, colour = model),
                  size = 0.5, linewidth = 0.5) +
  geom_text(aes(x = hi, label = label), hjust = -0.08, size = 2.7,
            colour = jama_black, show.legend = FALSE) +
  scale_x_log10(breaks = c(0.3, 0.5, 0.7, 1, 1.5, 2, 3)) +
  scale_y_continuous(breaks = 1:6, labels = rev(c("OS", "NRM", "aGVHD II\u2013IV", "CMV", "BSI", "IFI"))) +
  scale_shape_manual(values = c("M1" = 16, "M2" = 21), labels = c("M1 (Direct)", "M2 (Steroid-adjusted)")) +
  scale_colour_manual(values = c("M1" = jama_navy, "M2" = jama_steel),
                      labels = c("M1 (Direct)", "M2 (Steroid-adjusted)")) +
  scale_fill_manual(values = c("M1" = jama_navy, "M2" = "white"),
                    labels = c("M1 (Direct)", "M2 (Steroid-adjusted)")) +
  facet_wrap(~ comparison) +
  coord_cartesian(xlim = c(0.25, 5.5)) +
  labs(x = "Odds Ratio (log scale)", y = NULL,
       shape = NULL, colour = NULL, fill = NULL,
       caption = "Bayesian random-effects models (binomial family, brms). BSI C2 estimate is Mantel-Haenszel (not shown).") +
  theme_jama(base_size = 11) +
  theme(legend.position = "bottom")
```

### Key Findings

::: {.key-finding}
**1. Survival and GVHD benefit.** PTCy was associated with reduced overall mortality and acute GVHD in both Comparison 1 (vs CNI+MTX/MMF) and Comparison 2 (vs ATG). The effect was consistent in magnitude and direction across comparators.
:::

::: {.key-finding}
**2. NRM is steroid-mediated.** The apparent NRM benefit was substantially or completely attenuated after adjusting for GVHD-driven steroid exposure in both comparisons, indicating that PTCy reduces NRM indirectly — through fewer GVHD episodes and less corticosteroid toxicity.
:::

::: {.key-finding}
**3. CMV direction reversal.** PTCy increased CMV reactivation vs CNI+MTX/MMF (OR 1.41) but *decreased* it vs ATG (OR 0.77). This reversal confirms that CMV risk is ordered by the depth of T-cell depletion (CNI < PTCy < ATG), not by any intrinsic PTCy property.
:::

::: {.key-finding}
**4. BSI elevated in both comparisons.** Bloodstream infection risk was approximately doubled with PTCy regardless of comparator (C1: OR 2.39; C2: MH OR 2.03), though the C2 estimate is based on only 3 studies with heterogeneous definitions.
:::

### Summary of Evidence

```{r}
#| label: tbl-exec-summary
#| tbl-cap: "Summary of primary outcomes across both comparisons."

tibble::tribble(
  ~Outcome, ~Comparison, ~k, ~N, ~`OR [95% CrI]`, ~`P(directional)`, ~Certainty,
  "OS",            "C1", 35, 17661, "0.77 [0.71\u20130.83]", "P(<1) = 100%",  "\u2295\u2295\u2296\u2296 LOW",
  "OS",            "C2", 10, 10624, "0.69 [0.61\u20130.78]", "P(<1) = 100%",  "\u2295\u2295\u2296\u2296 LOW",
  "NRM",           "C1", 15,  2081, "0.86 [0.70\u20131.08]", "P(<1) = 90%",   "\u2295\u2296\u2296\u2296 V.LOW",
  "NRM",           "C2",  5,  1186, "0.65 [0.43\u20130.97]", "P(<1) = 98%",   "\u2295\u2296\u2296\u2296 V.LOW",
  "aGVHD II\u2013IV", "C1", 24,  3837, "0.63 [0.54\u20130.74]", "P(<1) = 100%",  "\u2295\u2295\u2296\u2296 LOW",
  "aGVHD II\u2013IV", "C2",  6,   784, "0.59 [0.42\u20130.84]", "P(<1) = 100%",  "\u2295\u2295\u2296\u2296 LOW",
  "CMV",           "C1", 18,  3357, "1.41 [1.20\u20131.65]", "P(>1) = 100%",  "\u2295\u2295\u2296\u2296 LOW",
  "CMV",           "C2", 12,  1267, "0.77 [0.60\u20130.99]", "P(<1) = 98%",   "\u2295\u2296\u2296\u2296 V.LOW",
  "BSI",           "C1",  6,  1319, "2.39 [1.85\u20133.10]", "P(>1) = 100%",  "\u2295\u2296\u2296\u2296 V.LOW",
  "BSI",           "C2",  3,   613, "2.03 [1.33\u20133.09]\u2020", "p = 0.001", "(\u2295\u2296\u2296\u2296)",
  "IFI",           "C1",  8,  3816, "0.73 [0.53\u20131.00]", "P(<1) = 97.5%", "\u2295\u2296\u2296\u2296 V.LOW",
  "IFI",           "C2", NA,    NA, "\u2014",                 "\u2014",         "\u2014"
) |>
  kable(align = c("l", "c", "r", "r", "c", "c", "c")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) |>
  footnote(symbol = c("Mantel-Haenszel fixed-effect estimate; formal GRADE not performed at k\u00a0=\u00a03."))
```

### Mechanistic Framework

The findings resolve into a unified framework built on two separable consequences of PTCy:

1. **GVHD suppression pathway** (intended): PTCy → reduced aGVHD → less steroid exposure → lower NRM → improved OS. The NRM benefit is fully steroid-mediated; the OS benefit is only partially mediated, suggesting an additional direct contribution.

2. **T-cell lymphodepletion pathway** (unintended): PTCy depletes CMV-specific and bacterial-specific T cells during engraftment. The CMV direction reversal across comparisons confirms this is a depth-of-depletion effect rather than a PTCy-specific one.

```{r}
#| label: fig-mech-diagram
#| fig-cap: "Mechanistic framework. The GVHD suppression pathway (left) mediates survival benefits; the T-cell depletion pathway (right) mediates infection risks. The CMV direction reversal across comparisons (inset) confirms that infection risk tracks depth of T-cell depletion."
#| echo: false
#| eval: false

# Placeholder — Mermaid diagram rendered by Quarto natively below
```

```{mermaid}
%%| fig-width: 8
%%| label: fig-mechanism
%%| fig-cap: "Dual-pathway mechanistic framework for PTCy effects."

flowchart LR
    PTCy --> A["↓ Alloreactive\nT cells"]
    PTCy --> B["↓ Pathogen-specific\nT cells"]

    A --> C["↓ aGVHD\nOR 0.59–0.63"]
    C --> D["↓ Steroid\nexposure"]
    D --> E["↓ NRM\n(fully mediated)"]
    D --> F["↓ OS\n(partially mediated)"]

    B --> G["↑ CMV vs CNI\nOR 1.41"]
    B --> H["↓ CMV vs ATG\nOR 0.77"]
    B --> I["↑ BSI\nOR 2.0–2.4"]

    style A fill:#a3c1d4,stroke:#2b5c8a,color:#1a2a3a
    style B fill:#d4d8dc,stroke:#6b7d8d,color:#1a2a3a
    style C fill:#a3c1d4,stroke:#2b5c8a,color:#1a2a3a
    style G fill:#d4d8dc,stroke:#6b7d8d,color:#1a2a3a
    style H fill:#d4d8dc,stroke:#6b7d8d,color:#1a2a3a
    style I fill:#d4d8dc,stroke:#6b7d8d,color:#1a2a3a
```

---

*Navigate to individual pages using the menu above for detailed results, model diagnostics, sensitivity analyses, GRADE assessments, and risk-of-bias evaluation.*
