---
title: "Post-Transplant Cyclophosphamide Meta-Analysis"
subtitle: "Bayesian random-effects meta-analysis of PTCy vs alternative GVHD prophylaxis platforms"
date: 2026-06-01
---

```{r}
#| label: setup
#| include: false
source("_common.R")

# Load all posterior draws
post_os      <- load_post("post_os.rds")
post_m2_os   <- load_post("post_m2_os.rds")
post_nrm     <- load_post("post_nrm.rds")
post_m2_nrm  <- load_post("post_m2_nrm.rds")
post_agvhd   <- load_post("post_agvhd.rds")
post_m1      <- load_post("post_m1.rds")
post_m2      <- load_post("post_m2.rds")
post_bsi     <- load_post("post_bsi.rds")
post_ifi     <- load_post("post_ifi.rds")

post_c2_os     <- load_post("post_c2_os.rds")
post_c2_m2_os  <- load_post("post_c2_m2_os.rds")
post_c2_nrm    <- load_post("post_c2_nrm.rds")
post_c2_agvhd  <- load_post("post_c2_agvhd.rds")
post_c2_cmv    <- load_post("post_c2_cmv.rds")
post_c2_m2_cmv <- load_post("post_c2_m2_cmv.rds")

# New outcomes: RRM, BK virus
post_rrm       <- tryCatch(load_post("post_c1_rrm.rds"), error = function(e) NULL)
post_c2_rrm    <- tryCatch(load_post("post_c2_rrm.rds"), error = function(e) NULL)
post_bk        <- tryCatch(load_post("post_c1_bk.rds"), error = function(e) NULL)
post_c2_bk     <- tryCatch(load_post("post_c2_bk.rds"), error = function(e) NULL)
post_irm       <- tryCatch(load_post("post_c1_irm.rds"), error = function(e) NULL)
post_cgvhd     <- tryCatch(load_post("post_c1_cgvhd_ms.rds"), error = function(e) NULL)
post_c2_cgvhd  <- tryCatch(load_post("post_c2_cgvhd_ms.rds"), error = function(e) NULL)
post_c2_irm    <- tryCatch(load_post("post_c2_irm.rds"), error = function(e) NULL)

# Dynamically compute OR summaries
or_os      <- or_summary(post_os,    "b_ptcy_binary", "OS", "C1: PTCy vs CNI+MTX")
or_m2_os   <- or_summary(post_m2_os, "b_ptcy_binary", "OS", "C1: PTCy vs CNI+MTX", "M2")
or_nrm     <- or_summary(post_nrm,    "b_ptcy_binary", "NRM", "C1: PTCy vs CNI+MTX")
or_m2_nrm  <- or_summary(post_m2_nrm, "b_ptcy_binary", "NRM", "C1: PTCy vs CNI+MTX", "M2")
or_agvhd   <- or_summary(post_agvhd, "b_ptcy_binary", "aGVHD II\u2013IV", "C1: PTCy vs CNI+MTX")
or_cmv     <- or_summary(post_m1,    "b_ptcy_binary", "CMV", "C1: PTCy vs CNI+MTX")
or_m2_cmv  <- or_summary(post_m2,    "b_ptcy_binary", "CMV", "C1: PTCy vs CNI+MTX", "M2")
or_bsi     <- or_summary(post_bsi,   "b_ptcy_binary", "BSI", "C1: PTCy vs CNI+MTX")
or_ifi     <- or_summary(post_ifi,   "b_ptcy_binary", "IFI", "C1: PTCy vs CNI+MTX")

or_c2_os     <- or_summary(post_c2_os,     "b_ptcy_binary", "OS", "C2: PTCy vs ATG")
or_c2_m2_os  <- or_summary(post_c2_m2_os,  "b_ptcy_binary", "OS", "C2: PTCy vs ATG", "M2")
or_c2_nrm    <- or_summary(post_c2_nrm,    "b_ptcy_binary", "NRM", "C2: PTCy vs ATG")
or_c2_agvhd  <- or_summary(post_c2_agvhd,  "b_ptcy_binary", "aGVHD II\u2013IV", "C2: PTCy vs ATG")
or_c2_cmv    <- or_summary(post_c2_cmv,    "b_ptcy_binary", "CMV", "C2: PTCy vs ATG")
or_c2_m2_cmv <- or_summary(post_c2_m2_cmv, "b_ptcy_binary", "CMV", "C2: PTCy vs ATG", "M2")

or_rrm       <- if (!is.null(post_rrm))    or_summary(post_rrm,    "b_ptcy_binary", "RRM", "C1: PTCy vs CNI+MTX") else list(or = NA, lo = NA, hi = NA)
or_c2_rrm    <- if (!is.null(post_c2_rrm)) or_summary(post_c2_rrm, "b_ptcy_binary", "RRM", "C2: PTCy vs ATG")     else list(or = NA, lo = NA, hi = NA)
or_bk        <- if (!is.null(post_bk))     or_summary(post_bk,     "b_ptcy_binary", "BK Virus", "C1: PTCy vs CNI+MTX") else list(or = NA, lo = NA, hi = NA)
or_c2_bk     <- if (!is.null(post_c2_bk))  or_summary(post_c2_bk,  "b_ptcy_binary", "BK Virus", "C2: PTCy vs ATG")     else list(or = NA, lo = NA, hi = NA)
or_irm       <- if (!is.null(post_irm))     or_summary(post_irm,     "b_ptcy_binary", "IRM", "C1: PTCy vs CNI+MTX")          else list(or = NA, lo = NA, hi = NA)
or_cgvhd     <- if (!is.null(post_cgvhd))    or_summary(post_cgvhd,    "b_ptcy_binary", "cGVHD mod-sev", "C1: PTCy vs CNI+MTX") else list(or = NA, lo = NA, hi = NA)
or_c2_cgvhd  <- if (!is.null(post_c2_cgvhd)) or_summary(post_c2_cgvhd, "b_ptcy_binary", "cGVHD mod-sev", "C2: PTCy vs ATG")     else list(or = NA, lo = NA, hi = NA)
or_c2_irm    <- if (!is.null(post_c2_irm)) or_summary(post_c2_irm, "b_ptcy_binary", "IRM", "C2: PTCy vs ATG")     else list(or = NA, lo = NA, hi = NA)

# Load Table2 for k/N values
t2_file <- file.path(model_dir, "Table2_post_block9.csv")
if (!file.exists(t2_file)) t2_file <- file.path("../03_models", "Table2_results.csv")
t2 <- tryCatch(read_csv(t2_file, show_col_types = FALSE), error = function(e) NULL)

# Database counts
studies_df <- read_csv(file.path(extract_dir, "studies.csv"), show_col_types = FALSE)
arms_df    <- read_csv(file.path(extract_dir, "arms.csv"), show_col_types = FALSE)
n_studies <- nrow(studies_df)
n_arms    <- nrow(arms_df)
n_rcts    <- sum(studies_df$study_design == "RCT", na.rm = TRUE)
```

## Abstract

::: {.callout-note appearance="minimal"}

**Background.** Post-transplant cyclophosphamide (PTCy) has become widely adopted for GVHD prophylaxis in allogeneic haematopoietic cell transplantation, yet comprehensive quantitative evidence synthesis across survival, GVHD, and infection outcomes — and across comparator regimens — has been lacking.

**Methods.** We conducted a Bayesian random-effects meta-analysis of `r n_studies` studies (`r n_rcts` RCTs) comparing PTCy to (1) CNI+MTX/MMF-based regimens (Comparison 1), (2) ATG-based regimens (Comparison 2), and (3) within-PTCy regimen variants (Comparison 3). Ten outcomes were analysed: OS, NRM, relapse-related mortality (RRM), acute GVHD grade II–IV, chronic GVHD moderate-severe (NIH), CMV reactivation, BK virus, BSI, IFI, and infection-related mortality (IRM). Steroid-mediation models (M2) tested whether GVHD-driven steroid exposure confounded observed effects.

**Results.** PTCy was associated with improved OS (C1: OR `r fmt_or(or_os$or, or_os$lo, or_os$hi)`; C2: OR `r fmt_or(or_c2_os$or, or_c2_os$lo, or_c2_os$hi)`), reduced aGVHD (C1: OR `r round(or_agvhd$or, 2)`; C2: OR `r round(or_c2_agvhd$or, 2)`), and a substantial reduction in moderate-severe chronic GVHD (C1: OR `r round(or_cgvhd$or, 2)` — the largest treatment effect). RRM was reduced in both comparisons (C1: OR `r round(or_rrm$or, 2)`; C2: OR `r round(or_c2_rrm$or, 2)`), supporting preserved GVL activity. CMV reactivation was increased vs CNI+MTX/MMF (OR `r fmt_or(or_cmv$or, or_cmv$lo, or_cmv$hi)`) but near-null vs ATG, consistent with a T-cell depletion depth gradient. BK virus (C1: OR `r round(or_bk$or, 2)`) and BSI (C1: OR `r round(or_bsi$or, 2)`) were elevated regardless of comparator. Despite elevated infection rates, infection-related mortality was not increased (C1: OR `r round(or_irm$or, 2)`, Bayesian–frequentist discordant; C2: OR `r round(or_c2_irm$or, 2)`). Within-PTCy regimen variants (C3) showed no significant differences.

**Conclusions.** PTCy provides consistent survival, GVHD, and anti-relapse advantages regardless of comparator. Chronic GVHD reduction is the largest treatment effect (OR 0.33). Infection rates are elevated but infection-related mortality is not clearly increased, suggesting the infectious cost is manageable. GRADE certainty is LOW for OS, aGVHD, and RRM; VERY LOW for other outcomes.
:::

---

## Executive Summary

```{r}
#| label: fig-exec-summary
#| fig-cap: "Bayesian meta-analysis — PTCy effect across outcomes and comparators. Filled circles represent direct-effect models (M1); open circles represent steroid-adjusted models (M2). The dashed vertical line marks OR = 1 (no effect). OR < 1 favours PTCy for mortality, GVHD, and IFI outcomes."
#| fig-height: 8
#| fig-width: 10

exec_list <- list(
  or_os, or_m2_os, or_nrm, or_m2_nrm, or_agvhd,
  or_cmv, or_m2_cmv, or_bsi, or_ifi,
  or_c2_os, or_c2_m2_os, or_c2_nrm, or_c2_agvhd,
  or_c2_cmv, or_c2_m2_cmv
)
# Add new outcomes if available
if (!is.null(post_rrm))    exec_list <- c(exec_list, list(or_rrm))
if (!is.null(post_c2_rrm)) exec_list <- c(exec_list, list(or_c2_rrm))
if (!is.null(post_bk))     exec_list <- c(exec_list, list(or_bk))
if (!is.null(post_c2_bk))  exec_list <- c(exec_list, list(or_c2_bk))
if (!is.null(post_cgvhd))    exec_list <- c(exec_list, list(or_cgvhd))
if (!is.null(post_c2_cgvhd)) exec_list <- c(exec_list, list(or_c2_cgvhd))
if (!is.null(post_irm))    exec_list <- c(exec_list, list(or_irm))
if (!is.null(post_c2_irm)) exec_list <- c(exec_list, list(or_c2_irm))

outcome_levels <- rev(c("OS", "NRM", "RRM", "aGVHD II\u2013IV", "cGVHD mod-sev", "CMV", "BK Virus", "BSI", "IFI", "IRM"))

pd <- do.call(rbind, exec_list) |>
  as_tibble() |>
  mutate(
    outcome = factor(outcome, levels = outcome_levels),
    comparison = factor(comparison, levels = c("C1: PTCy vs CNI+MTX", "C2: PTCy vs ATG")),
    label = sprintf("%.2f [%.2f, %.2f]", or, lo, hi),
    y_num = as.numeric(outcome),
    y_pos = y_num + ifelse(model == "M2", -0.15, 0.15)
  )

n_outcomes <- length(unique(pd$outcome))

ggplot(pd, aes(x = or, y = y_pos, shape = model, fill = model)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = jama_grey, linewidth = 0.4) +
  geom_pointrange(aes(xmin = lo, xmax = hi, colour = model),
                  size = 0.5, linewidth = 0.5) +
  geom_text(aes(x = hi, label = label), hjust = -0.05, size = 2.5,
            colour = jama_black, show.legend = FALSE, nudge_y = 0.25) +
  scale_x_log10(breaks = c(0.3, 0.5, 0.7, 1, 1.5, 2, 3)) +
  scale_y_continuous(breaks = seq_len(n_outcomes), labels = outcome_levels[seq_len(n_outcomes)]) +
  scale_shape_manual(values = c("M1" = 16, "M2" = 21), labels = c("M1 (Direct)", "M2 (Steroid-adjusted)")) +
  scale_colour_manual(values = c("M1" = jama_navy, "M2" = jama_steel),
                      labels = c("M1 (Direct)", "M2 (Steroid-adjusted)")) +
  scale_fill_manual(values = c("M1" = jama_navy, "M2" = "white"),
                    labels = c("M1 (Direct)", "M2 (Steroid-adjusted)")) +
  facet_wrap(~ comparison) +
  coord_cartesian(xlim = c(0.25, 5.5)) +
  labs(x = "Odds Ratio (log scale)", y = NULL,
       shape = NULL, colour = NULL, fill = NULL,
       caption = "Bayesian random-effects models (binomial family, brms). Post-Block-9 update.") +
  theme_jama(base_size = 11) +
  theme(legend.position = "bottom")
```

### Key Findings

::: {.key-finding}
**1. Chronic GVHD — the largest treatment effect.** PTCy reduced moderate-severe chronic GVHD by two-thirds vs CNI+MTX (C1 OR `r round(or_cgvhd$or, 2)` [95% CrI `r round(or_cgvhd$lo, 2)`–`r round(or_cgvhd$hi, 2)`], k = 21). This is the largest treatment effect in the analysis — substantially exceeding acute GVHD (OR `r round(or_agvhd$or, 2)`) — with direct implications for long-term quality of life and immunosuppression duration. The effect attenuates vs ATG (C2 OR `r round(or_c2_cgvhd$or, 2)`), consistent with ATG also providing cGVHD prophylaxis.
:::

::: {.key-finding}
**2. Survival and acute GVHD benefit.** PTCy was associated with reduced overall mortality (C1 OR `r round(or_os$or, 2)`; C2 OR `r round(or_c2_os$or, 2)`) and acute GVHD grade II–IV (C1 OR `r round(or_agvhd$or, 2)`; C2 OR `r round(or_c2_agvhd$or, 2)`) in both comparisons. Effects were consistent across comparators.
:::

::: {.key-finding}
**3. Anti-relapse benefit preserved.** RRM was reduced in both comparisons (C1 OR `r round(or_rrm$or, 2)`; C2 OR `r round(or_c2_rrm$or, 2)`) with no steroid mediation, supporting preserved graft-versus-leukaemia activity.
:::

::: {.key-finding}
**4. Infections increased but infection-related mortality not clearly elevated.** CMV, BSI, and BK virus rates were elevated vs CNI+MTX. However, infection-related mortality was not clearly increased (C1 Bayesian OR `r round(or_irm$or, 2)` [Bayesian–frequentist discordant]; C2 OR `r round(or_c2_irm$or, 2)`). The infectious morbidity appears manageable and does not translate proportionally into death.
:::

::: {.key-finding}
**5. CMV depletion gradient.** PTCy increased CMV vs CNI+MTX (OR `r round(or_cmv$or, 2)`) but was near-null vs ATG (OR `r round(or_c2_cmv$or, 2)`), supporting a T-cell depletion depth gradient (CNI < PTCy ≈ ATG) rather than a PTCy-specific effect.
:::

::: {.key-finding}
**6. Within-PTCy variants equivalent.** C3 found no significant differences across PTCy regimen variants for any outcome, suggesting the PTCy component dominates the regimen effect.
:::

### Summary of Evidence

```{r}
#| label: tbl-exec-summary
#| tbl-cap: "Summary of primary outcomes across both comparisons."

# Build dynamically from posterior draws
get_k <- function(slug, mdl = "m1") {
  if (!is.null(t2)) {
    row <- t2 |> filter(slug == !!slug, model == !!mdl)
    if (nrow(row) > 0) return(list(k = row$k, n = row$n_total))
  }
  list(k = NA, n = NA)
}

summary_rows <- list(
  list("OS",            "C1", fmt_or(or_os$or, or_os$lo, or_os$hi),               sprintf("P(<1) = %.0f%%", or_os$p_dir * 100)),
  list("OS",            "C2", fmt_or(or_c2_os$or, or_c2_os$lo, or_c2_os$hi),       sprintf("P(<1) = %.0f%%", or_c2_os$p_dir * 100)),
  list("NRM",           "C1", fmt_or(or_nrm$or, or_nrm$lo, or_nrm$hi),             sprintf("P(<1) = %.0f%%", or_nrm$p_dir * 100)),
  list("aGVHD II\u2013IV", "C1", fmt_or(or_agvhd$or, or_agvhd$lo, or_agvhd$hi),   sprintf("P(<1) = %.0f%%", or_agvhd$p_dir * 100)),
  list("aGVHD II\u2013IV", "C2", fmt_or(or_c2_agvhd$or, or_c2_agvhd$lo, or_c2_agvhd$hi), sprintf("P(<1) = %.0f%%", or_c2_agvhd$p_dir * 100)),
  list("CMV",           "C1", fmt_or(or_cmv$or, or_cmv$lo, or_cmv$hi),             sprintf("P(>1) = %.0f%%", or_cmv$p_dir * 100)),
  list("CMV",           "C2", fmt_or(or_c2_cmv$or, or_c2_cmv$lo, or_c2_cmv$hi),   sprintf("P(<1) = %.0f%%", or_c2_cmv$p_dir * 100)),
  list("BSI",           "C1", fmt_or(or_bsi$or, or_bsi$lo, or_bsi$hi),             sprintf("P(>1) = %.0f%%", or_bsi$p_dir * 100)),
  list("IFI",           "C1", fmt_or(or_ifi$or, or_ifi$lo, or_ifi$hi),             sprintf("P(<1) = %.0f%%", or_ifi$p_dir * 100))
)
if (!is.null(post_rrm)) {
  summary_rows <- c(summary_rows, list(
    list("RRM", "C1", fmt_or(or_rrm$or, or_rrm$lo, or_rrm$hi), sprintf("P(<1) = %.0f%%", or_rrm$p_dir * 100))
  ))
}
if (!is.null(post_c2_rrm)) {
  summary_rows <- c(summary_rows, list(
    list("RRM", "C2", fmt_or(or_c2_rrm$or, or_c2_rrm$lo, or_c2_rrm$hi), sprintf("P(<1) = %.0f%%", or_c2_rrm$p_dir * 100))
  ))
}
if (!is.null(post_bk)) {
  summary_rows <- c(summary_rows, list(
    list("BK Virus", "C1", fmt_or(or_bk$or, or_bk$lo, or_bk$hi), sprintf("P(>1) = %.0f%%", or_bk$p_dir * 100))
  ))
}
if (!is.null(post_c2_bk)) {
  summary_rows <- c(summary_rows, list(
    list("BK Virus", "C2", fmt_or(or_c2_bk$or, or_c2_bk$lo, or_c2_bk$hi), sprintf("P(>1) = %.0f%%", or_c2_bk$p_dir * 100))
  ))
}
if (!is.null(post_cgvhd)) {
  summary_rows <- c(summary_rows, list(
    list("cGVHD mod-sev", "C1", fmt_or(or_cgvhd$or, or_cgvhd$lo, or_cgvhd$hi), sprintf("P(<1) = %.0f%%", or_cgvhd$p_dir * 100))
  ))
}
if (!is.null(post_c2_cgvhd)) {
  summary_rows <- c(summary_rows, list(
    list("cGVHD mod-sev", "C2", fmt_or(or_c2_cgvhd$or, or_c2_cgvhd$lo, or_c2_cgvhd$hi), sprintf("P(<1) = %.0f%%", or_c2_cgvhd$p_dir * 100))
  ))
}
if (!is.null(post_irm)) {
  summary_rows <- c(summary_rows, list(
    list("IRM", "C1", fmt_or(or_irm$or, or_irm$lo, or_irm$hi), sprintf("P(>1) = %.0f%%", or_irm$p_dir * 100))
  ))
}
if (!is.null(post_c2_irm)) {
  summary_rows <- c(summary_rows, list(
    list("IRM", "C2", fmt_or(or_c2_irm$or, or_c2_irm$lo, or_c2_irm$hi), sprintf("P(<1) = %.0f%%", or_c2_irm$p_dir * 100))
  ))
}

summary_tbl <- do.call(rbind, lapply(summary_rows, function(r) {
  tibble(Outcome = r[[1]], Comparison = r[[2]], `OR [95% CrI]` = r[[3]], `P(directional)` = r[[4]])
}))

summary_tbl |>
  kable(align = c("l", "c", "c", "c")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

### Mechanistic Framework

The findings resolve into a framework built on three separable consequences of PTCy:

1. **GVHD suppression pathway** (intended): PTCy eliminates alloreactive T cells → reduced aGVHD (OR 0.67) and profoundly reduced cGVHD (OR 0.33) → less steroid exposure → lower NRM → improved OS. Chronic GVHD is the most affected outcome, consistent with PTCy's early alloreactive depletion having an outsized impact on chronic alloimmune processes.

2. **GVL preservation pathway** (intended): PTCy → reduced relapse-related mortality (C1 OR 0.84; C2 OR 0.79), steroid-independent. This explains the residual OS advantage after GVHD/steroid adjustment.

3. **T-cell lymphodepletion pathway** (unintended): PTCy depletes pathogen-specific T cells during engraftment → elevated CMV, BSI, and BK virus. The CMV depletion gradient (CNI < PTCy ≈ ATG) confirms this is a relative depth-of-depletion effect. BK virus has an additional urothelial toxicity component. Critically, the elevated infection *rates* do not translate clearly into elevated infection-related *mortality*, suggesting the infectious cost is clinically manageable.

```{mermaid}
%%| fig-width: 9
%%| label: fig-mechanism
%%| fig-cap: "Three-pathway mechanistic framework for PTCy effects."

flowchart LR
    PTCy --> A["↓ Alloreactive\nT cells"]
    PTCy --> B["↓ Pathogen-specific\nT cells"]
    PTCy --> U["Urothelial\ntoxicity"]

    A --> C["↓↓ cGVHD\nOR 0.33"]
    A --> C2["↓ aGVHD\nOR 0.67"]
    A --> R["↓ RRM\nOR 0.84"]
    C --> D["↓ Steroid\nexposure"]
    C2 --> D
    D --> E["↓ NRM"]
    D --> F["↓ OS\nOR 0.79"]
    R --> F

    B --> G["↑ CMV vs CNI\nOR 1.26"]
    B --> H["≈ CMV vs ATG\nOR 0.92"]
    B --> I["↑ BSI\nOR 1.87"]
    B --> K["IRM ≈ null\nOR 1.01–1.35"]
    B --> J["↑ BK Virus"]
    U --> J["↑ BK Virus\nOR 2.48"]

    style A fill:#a3c1d4,stroke:#2b5c8a,color:#1a2a3a
    style B fill:#d4d8dc,stroke:#6b7d8d,color:#1a2a3a
    style U fill:#d4d8dc,stroke:#6b7d8d,color:#1a2a3a
    style C fill:#a3c1d4,stroke:#2b5c8a,color:#1a2a3a
    style C2 fill:#a3c1d4,stroke:#2b5c8a,color:#1a2a3a
    style R fill:#a3c1d4,stroke:#2b5c8a,color:#1a2a3a
    style G fill:#d4d8dc,stroke:#6b7d8d,color:#1a2a3a
    style H fill:#d4d8dc,stroke:#6b7d8d,color:#1a2a3a
    style I fill:#d4d8dc,stroke:#6b7d8d,color:#1a2a3a
    style J fill:#d4d8dc,stroke:#6b7d8d,color:#1a2a3a
    style K fill:#d4d8dc,stroke:#6b7d8d,color:#1a2a3a
```

---

*Navigate to individual pages using the menu above for detailed results, model diagnostics, sensitivity analyses, GRADE assessments, and risk-of-bias evaluation.*

::: {.callout-tip}
## Post-Block-9 Update (`r Sys.Date()`)
This analysis includes `r n_studies` studies (`r n_arms` arms, `r n_rcts` RCTs) extracted through Block 9, an expansion from the original 220-study baseline. Models loading from: **`r basename(model_dir)`**.
:::
