---
title: "Comparison 2: PTCy vs ATG"
---

```{r}
#| label: setup
#| include: false
source("_common.R")

post_c2_os      <- load_post("post_c2_os.rds")
post_c2_m2_os   <- load_post("post_c2_m2_os.rds")
post_c2_nrm     <- load_post("post_c2_nrm.rds")
post_c2_m2_nrm  <- tryCatch(load_post("post_c2_m2_nrm.rds"), error = function(e) NULL)
post_c2_agvhd   <- load_post("post_c2_agvhd.rds")
post_c2_cmv     <- load_post("post_c2_cmv.rds")
post_c2_m2_cmv  <- load_post("post_c2_m2_cmv.rds")
post_c2_rrm     <- tryCatch(load_post("post_c2_rrm.rds"), error = function(e) NULL)
post_c2_bk      <- tryCatch(load_post("post_c2_bk.rds"), error = function(e) NULL)
post_c2_irm     <- tryCatch(load_post("post_c2_irm.rds"), error = function(e) NULL)
post_c2_cgvhd   <- tryCatch(load_post("post_c2_cgvhd_ms.rds"), error = function(e) NULL)

# C1 posteriors for cross-comparison
post_m1 <- load_post("post_m1.rds")
post_os <- load_post("post_os.rds")
```

Comparison 2 evaluates PTCy against ATG-based GVHD prophylaxis — a more deeply T-cell–depleting comparator. This comparison draws on paired studies after excluding PTCy+ATG combination, mixed, and unpaired arms.

## Overall Survival {#sec-c2-os}

```{r}
#| label: fig-c2-os
#| fig-cap: "OS posterior densities for Comparison 2. M2 steroid adjustment attenuates the effect by 35%, paralleling the C1 pattern."
#| fig-height: 4.5

tibble(
  M1 = post_c2_os$b_ptcy_binary,
  M2 = if ("b_ptcy_binary" %in% names(post_c2_m2_os)) post_c2_m2_os$b_ptcy_binary else post_c2_m2_os$b_eventsn_ptcy_binary
) |>
  pivot_longer(everything(), names_to = "Model", values_to = "log_or") |>
  ggplot(aes(x = log_or, fill = Model, colour = Model)) +
  geom_density(alpha = 0.35, linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = jama_grey) +
  scale_fill_manual(values = c("M1" = jama_navy, "M2" = jama_steel)) +
  scale_colour_manual(values = c("M1" = jama_navy, "M2" = jama_steel)) +
  labs(x = "log-OR (PTCy vs ATG)", y = "Posterior density",
       title = "Overall Survival — Comparison 2",
       subtitle = "k = 9, N = 12,791 · M1 OR 0.83 → M2 OR 0.81 (minimal attenuation)") +
  theme_jama()
```

**M1:** OR 0.83 (95% CrI 0.75–0.92), P(OR < 1) = 100%, τ = 0.50, k = 9. Consistent OS advantage for PTCy over ATG, though attenuated from the pre-Block-9 estimate (0.69).

**M2:** OR 0.82 (95% CrI 0.71–0.94), minimal log-OR attenuation (~6%). Unlike the pre-Block-9 analysis (35% attenuation), steroid adjustment barely changes the estimate, suggesting that the OS benefit in C2 operates largely through non-GVHD pathways.

## Non-Relapse Mortality {#sec-c2-nrm}

**M1:** k reduced to 3 after arm reclassification (from 5 pre-Block-9). The frequentist estimate is OR 0.63 (95% CI 0.30–1.34), I² = 0%. With only 3 studies, the CrI is too wide for reliable inference. M2 steroid-adjusted models were not fit due to insufficient k.

::: {.callout-warning}
## C2 NRM Below Modelling Threshold
The pre-Block-9 C2 NRM (k = 5, OR 0.65 [0.43–0.97]) was nominally significant. Arm reclassification corrections reduced k to 3, and the estimate should now be considered hypothesis-generating only.
:::

## Acute GVHD Grade II–IV {#sec-c2-agvhd}

**M1:** OR 0.63 (95% CrI 0.47–0.83), P(OR < 1) > 99%, τ = 0.53, k = 8. Concordant with C1 (OR 0.67), confirming aGVHD benefit is comparator-independent.

## Chronic GVHD Moderate-Severe {#sec-c2-cgvhd}

```{r}
#| label: fig-c2-cgvhd
#| fig-cap: "Posterior density for the PTCy effect on moderate-severe chronic GVHD (C2)."
#| fig-height: 4

if (!is.null(post_c2_cgvhd)) {
  tibble(log_or = post_c2_cgvhd$b_ptcy_binary) |>
    ggplot(aes(x = log_or)) +
    geom_density(fill = jama_navy, colour = jama_navy, alpha = 0.4, linewidth = 0.6) +
    geom_vline(xintercept = 0, linetype = "dashed", colour = jama_grey) +
    labs(x = "log-OR (PTCy vs ATG)", y = "Posterior density",
         title = "Chronic GVHD Moderate-Severe — Comparison 2",
         subtitle = "k = 6, N = 905 · OR 0.79 [0.55\u20131.14]") +
    theme_jama()
}
```

**M1:** OR 0.79 (95% CrI 0.55–1.14), P(OR < 1) = 90%, k = 6, τ = 0.89. Directionally protective but CrI crosses null. The C2 cGVHD effect (OR 0.79) is weaker than C1 (OR 0.47), consistent with ATG also providing significant cGVHD prophylaxis through T-cell depletion — the differential between PTCy and ATG for chronic GVHD is smaller than between PTCy and CNI+MTX.

## CMV Reactivation — The Direction Reversal {#sec-c2-cmv}

```{r}
#| label: fig-c2-cmv-reversal
#| fig-cap: "CMV posterior densities across comparisons. In C1 (navy), PTCy increases CMV risk (posterior right of null). In C2 (grey), PTCy decreases CMV risk (posterior left of null). This reversal confirms that CMV risk tracks the depth of T-cell depletion."
#| fig-height: 5

tibble(
  `C1: vs CNI+MTX` = post_m1$b_ptcy_binary,
  `C2: vs ATG` = post_c2_cmv$b_ptcy_binary
) |>
  pivot_longer(everything(), names_to = "Comparison", values_to = "log_or") |>
  ggplot(aes(x = log_or, fill = Comparison, colour = Comparison)) +
  geom_density(alpha = 0.3, linewidth = 0.7) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = jama_grey) +
  annotate("text", x = 0.45, y = 0.05, label = "PTCy\nharmful",
           colour = jama_steel, size = 3, fontface = "italic") +
  annotate("text", x = -0.45, y = 0.05, label = "PTCy\nprotective",
           colour = jama_steel, size = 3, fontface = "italic") +
  scale_fill_manual(values = c("C1: vs CNI+MTX" = jama_navy, "C2: vs ATG" = jama_steel)) +
  scale_colour_manual(values = c("C1: vs CNI+MTX" = jama_navy, "C2: vs ATG" = jama_steel)) +
  labs(x = "log-OR (PTCy vs comparator)", y = "Posterior density",
       title = "CMV Reactivation — Direction Reversal Across Comparisons",
       subtitle = "C1 OR 1.26 [1.07\u20131.47] vs C2 OR 0.97 [0.77\u20131.23]") +
  theme_jama()
```

::: {.callout-note}
## CMV: From Direction Reversal to Depletion Gradient
In the pre-Block-9 analysis, a qualitative direction reversal was observed (C1 OR 1.41 vs C2 OR 0.77 [0.60–0.99]). With the expanded post-Block-9 corpus (C1 k = 22, C2 k = 12), C1 CMV remains harmful (OR 1.26 [1.07–1.47]) but C2 CMV is null (OR 0.97 [0.77–1.23]). The evidence now supports a **depletion gradient** (C1 harmful → C2 near-null → C2-sensitivity null) rather than a binary reversal. See [Discussion](discussion.qmd) for the full three-point gradient interpretation.
:::

## BSI Descriptive Supplement {#sec-c2-bsi}

```{r}
#| label: fig-c2-bsi-forest
#| fig-cap: "BSI forest plot for Comparison 2 (k = 3). Substantial heterogeneity (I² = 69%) driven by definitional differences and a donor-type confound in Meyer 2025. The diamond represents the Mantel-Haenszel fixed-effect pooled estimate."
#| fig-height: 4

mh_bsi <- load_model("c2_mh_bsi.rds")

bsi_studies <- tibble(
  study = c("Meyer 2025", "Berro 2021", "Bordat 2026"),
  or = exp(mh_bsi$yi),
  lo = exp(mh_bsi$yi - 1.96 * sqrt(mh_bsi$vi)),
  hi = exp(mh_bsi$yi + 1.96 * sqrt(mh_bsi$vi)),
  weight = round((1/mh_bsi$vi) / sum(1/mh_bsi$vi) * 100, 1)
) |>
  mutate(label = sprintf("%.2f [%.2f, %.2f]  (%s%%)", or, lo, hi, weight))

pooled <- tibble(
  study = "MH Pooled", or = exp(mh_bsi$beta),
  lo = exp(mh_bsi$ci.lb), hi = exp(mh_bsi$ci.ub),
  weight = 100, label = sprintf("%.2f [%.2f, %.2f]", exp(mh_bsi$beta),
                                 exp(mh_bsi$ci.lb), exp(mh_bsi$ci.ub))
)

bsi_all <- bind_rows(bsi_studies, pooled) |>
  mutate(study = factor(study, levels = rev(c("Meyer 2025", "Berro 2021", "Bordat 2026", "MH Pooled"))),
         is_pooled = study == "MH Pooled")

ggplot(bsi_all, aes(x = or, y = study)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = jama_grey) +
  geom_pointrange(aes(xmin = lo, xmax = hi, shape = is_pooled, size = is_pooled),
                  colour = jama_navy, fill = jama_navy) +
  geom_text(aes(x = hi, label = label), hjust = -0.08, size = 3, colour = jama_black) +
  scale_shape_manual(values = c("FALSE" = 16, "TRUE" = 18), guide = "none") +
  scale_size_manual(values = c("FALSE" = 0.5, "TRUE" = 0.8), guide = "none") +
  scale_x_log10(breaks = c(0.5, 1, 2, 4)) +
  coord_cartesian(xlim = c(0.3, 8)) +
  labs(x = "Odds Ratio (log scale)", y = NULL,
       title = "BSI \u2014 Mantel-Haenszel Forest Plot (C2)",
       subtitle = sprintf("MH OR 2.03, I\u00B2 = 69%%, k = 3")) +
  theme_jama()
```

MH OR 2.03 (95% CI 1.33–3.09), but driven by definition mismatch (any culture vs ICU/fatal only), timepoint mismatch (D+100 vs 1 year), and donor-type confound (Meyer 2025: 54% haplo vs 100% MUD). Hypothesis-generating only.

## Relapse-Related Mortality {#sec-c2-rrm}

```{r}
#| label: fig-c2-rrm
#| fig-cap: "Posterior density for the PTCy effect on relapse-related mortality (C2)."
#| fig-height: 4


if (!is.null(post_c2_rrm)) {
  tibble(log_or = post_c2_rrm$b_ptcy_binary) |>
    ggplot(aes(x = log_or)) +
    geom_density(fill = jama_navy, colour = jama_navy, alpha = 0.4, linewidth = 0.6) +
    geom_vline(xintercept = 0, linetype = "dashed", colour = jama_grey) +
    labs(x = "log-OR (PTCy vs ATG)", y = "Posterior density",
         title = "Relapse-Related Mortality — Comparison 2",
         subtitle = "k = 10, N = 17,182 · OR 0.79 [0.69\u20130.89]") +
    theme_jama()
}
```

**M1:** OR 0.79 (95% CrI 0.69–0.89), P(OR < 1) > 99%, k = 10, τ = 0.14. A clear RRM advantage for PTCy over ATG, consistent with the C1 finding (OR 0.86) and supporting preserved GVL activity. Notably, the C2 RRM τ (0.14) is the lowest of any outcome, indicating minimal between-study heterogeneity.

## BK Virus Reactivation {#sec-c2-bk}

```{r}
#| label: fig-c2-bk
#| fig-cap: "Posterior density for the PTCy effect on BK virus reactivation (C2)."
#| fig-height: 4


if (!is.null(post_c2_bk)) {
  tibble(log_or = post_c2_bk$b_ptcy_binary) |>
    ggplot(aes(x = log_or)) +
    geom_density(fill = jama_navy, colour = jama_navy, alpha = 0.4, linewidth = 0.6) +
    geom_vline(xintercept = 0, linetype = "dashed", colour = jama_grey) +
    labs(x = "log-OR (PTCy vs ATG)", y = "Posterior density",
         title = "BK Virus Reactivation — Comparison 2",
         subtitle = "k = 3, N = 313 · OR 2.44 [1.40\u20134.26]") +
    theme_jama()
}
```

**M1:** OR 2.44 (95% CrI 1.40–4.26), P(OR > 1) > 99%, k = 3. Consistent with C1 (OR 2.48), confirming elevated BK virus risk is comparator-independent. This supports the dual-mechanism hypothesis (urothelial toxicity + T-cell depletion), since ATG provides comparable T-cell depletion but not urothelial damage.

## Infection-Related Mortality {#sec-c2-irm}

```{r}
#| label: fig-c2-irm
#| fig-cap: "Posterior density for the PTCy effect on infection-related mortality (C2)."
#| fig-height: 4

if (!is.null(post_c2_irm)) {
  tibble(log_or = post_c2_irm$b_ptcy_binary) |>
    ggplot(aes(x = log_or)) +
    geom_density(fill = jama_navy, colour = jama_navy, alpha = 0.4, linewidth = 0.6) +
    geom_vline(xintercept = 0, linetype = "dashed", colour = jama_grey) +
    labs(x = "log-OR (PTCy vs ATG)", y = "Posterior density",
         title = "Infection-Related Mortality — Comparison 2",
         subtitle = "k = 7, N = 13,346 · OR 0.90 [0.74\u20131.09]") +
    theme_jama()
}
```

**Bayesian M1:** OR 0.90 (95% CrI 0.74–1.09), P(OR < 1) = 86%, k = 7, τ = 0.58. Directionally protective (lower IRM with PTCy than ATG), concordant with the frequentist estimate (OR 0.81 [0.66–1.00], I² = 0%). Unlike C1, Bayesian and frequentist models agree: PTCy does not increase infection-related mortality relative to ATG.

::: {.callout-note}
## IRM Across Comparisons
The C1/C2 IRM pattern mirrors the CMV depletion gradient: IRM is not significantly elevated vs CNI+MTX (Bayesian OR 1.19 [1.00–1.43]) but *reduced* vs ATG (OR 0.90). This is consistent with ATG causing greater T-cell depletion–related mortality than PTCy. The C1 Bayesian–frequentist divergence warrants caution, but the overall picture suggests PTCy's infection cost is offset by ATG's even greater immune compromise.
:::

## Cross-Comparison Summary {#sec-c2-summary}

```{r}
#| label: fig-c2-combined
#| fig-cap: "Combined forest plot comparing Comparison 1 (left) and Comparison 2 (right). Post-Block-9 update with expanded outcomes including RRM and BK virus."
#| fig-height: 9
#| fig-width: 11

get_ptcy_col <- function(post) {
  if ("b_ptcy_binary" %in% names(post)) return("b_ptcy_binary")
  if ("b_eventsn_ptcy_binary" %in% names(post)) return("b_eventsn_ptcy_binary")
  stop("Cannot find ptcy_binary column in posterior")
}

c1_list <- list(
  or_summary(post_os,    "b_ptcy_binary", "OS", "C1: PTCy vs CNI+MTX"),
  or_summary(load_post("post_m2_os.rds"), get_ptcy_col(load_post("post_m2_os.rds")), "OS", "C1: PTCy vs CNI+MTX", "M2"),
  or_summary(load_post("post_nrm.rds"),    "b_ptcy_binary", "NRM", "C1: PTCy vs CNI+MTX"),
  or_summary(load_post("post_m2_nrm.rds"), get_ptcy_col(load_post("post_m2_nrm.rds")), "NRM", "C1: PTCy vs CNI+MTX", "M2"),
  or_summary(load_post("post_agvhd.rds"), "b_ptcy_binary", "aGVHD II\u2013IV", "C1: PTCy vs CNI+MTX"),
  or_summary(post_m1, "b_ptcy_binary", "CMV", "C1: PTCy vs CNI+MTX"),
  or_summary(load_post("post_m2.rds"), get_ptcy_col(load_post("post_m2.rds")), "CMV", "C1: PTCy vs CNI+MTX", "M2"),
  or_summary(load_post("post_bsi.rds"),   "b_ptcy_binary", "BSI", "C1: PTCy vs CNI+MTX"),
  or_summary(load_post("post_ifi.rds"),   "b_ptcy_binary", "IFI", "C1: PTCy vs CNI+MTX")
)
# Add RRM and BK for C1 if available
post_c1_rrm_tmp <- tryCatch(load_post("post_c1_rrm.rds"), error = function(e) NULL)
post_c1_bk_tmp  <- tryCatch(load_post("post_c1_bk.rds"), error = function(e) NULL)
if (!is.null(post_c1_rrm_tmp)) c1_list <- c(c1_list, list(or_summary(post_c1_rrm_tmp, "b_ptcy_binary", "RRM", "C1: PTCy vs CNI+MTX")))
if (!is.null(post_c1_bk_tmp))  c1_list <- c(c1_list, list(or_summary(post_c1_bk_tmp,  "b_ptcy_binary", "BK Virus", "C1: PTCy vs CNI+MTX")))

c2_list <- list(
  or_summary(post_c2_os,     "b_ptcy_binary", "OS", "C2: PTCy vs ATG"),
  or_summary(load_post("post_c2_m2_os.rds"), get_ptcy_col(load_post("post_c2_m2_os.rds")), "OS", "C2: PTCy vs ATG", "M2"),
  or_summary(post_c2_nrm,    "b_ptcy_binary", "NRM", "C2: PTCy vs ATG"),
  or_summary(post_c2_agvhd,  "b_ptcy_binary", "aGVHD II\u2013IV", "C2: PTCy vs ATG"),
  or_summary(post_c2_cmv,    "b_ptcy_binary", "CMV", "C2: PTCy vs ATG"),
  or_summary(load_post("post_c2_m2_cmv.rds"), get_ptcy_col(load_post("post_c2_m2_cmv.rds")), "CMV", "C2: PTCy vs ATG", "M2")
)
if (!is.null(post_c2_rrm)) c2_list <- c(c2_list, list(or_summary(post_c2_rrm, "b_ptcy_binary", "RRM", "C2: PTCy vs ATG")))
if (!is.null(post_c2_bk))  c2_list <- c(c2_list, list(or_summary(post_c2_bk,  "b_ptcy_binary", "BK Virus", "C2: PTCy vs ATG")))
if (!is.null(post_c2_irm)) c2_list <- c(c2_list, list(or_summary(post_c2_irm, "b_ptcy_binary", "IRM", "C2: PTCy vs ATG")))

post_c1_irm_tmp <- tryCatch(load_post("post_c1_irm.rds"), error = function(e) NULL)
if (!is.null(post_c1_irm_tmp)) c1_list <- c(c1_list, list(or_summary(post_c1_irm_tmp, "b_ptcy_binary", "IRM", "C1: PTCy vs CNI+MTX")))

post_c1_cgvhd_tmp <- tryCatch(load_post("post_c1_cgvhd_ms.rds"), error = function(e) NULL)
if (!is.null(post_c1_cgvhd_tmp)) c1_list <- c(c1_list, list(or_summary(post_c1_cgvhd_tmp, "b_ptcy_binary", "cGVHD mod-sev", "C1: PTCy vs CNI+MTX")))
if (!is.null(post_c2_cgvhd)) c2_list <- c(c2_list, list(or_summary(post_c2_cgvhd, "b_ptcy_binary", "cGVHD mod-sev", "C2: PTCy vs ATG")))

outcome_levels <- rev(c("OS", "NRM", "RRM", "aGVHD II\u2013IV", "cGVHD mod-sev", "CMV", "BK Virus", "BSI", "IFI", "IRM"))

all_pd <- do.call(rbind, c(c1_list, c2_list)) |>
  as_tibble() |>
  mutate(
    outcome = factor(outcome, levels = outcome_levels),
    comparison = factor(comparison, levels = c("C1: PTCy vs CNI+MTX", "C2: PTCy vs ATG")),
    label = sprintf("%.2f [%.2f, %.2f]", or, lo, hi),
    y_num = as.numeric(outcome),
    y_pos = y_num + ifelse(model == "M2", -0.15, 0.15)
  )

n_outcomes <- length(unique(all_pd$outcome))

ggplot(all_pd, aes(x = or, y = y_pos, shape = model, fill = model, colour = model)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = jama_grey) +
  geom_pointrange(aes(xmin = lo, xmax = hi), size = 0.45, linewidth = 0.5) +
  geom_text(aes(x = hi, label = label), hjust = -0.05, size = 2.5,
            colour = jama_black, show.legend = FALSE, nudge_y = 0.25) +
  scale_x_log10(breaks = c(0.3, 0.5, 0.7, 1, 1.5, 2, 3)) +
  scale_y_continuous(breaks = seq_len(n_outcomes),
                     labels = outcome_levels[seq_len(n_outcomes)]) +
  scale_shape_manual(values = c("M1" = 16, "M2" = 21)) +
  scale_colour_manual(values = c("M1" = jama_navy, "M2" = jama_steel)) +
  scale_fill_manual(values = c("M1" = jama_navy, "M2" = "white")) +
  facet_wrap(~ comparison) +
  coord_cartesian(xlim = c(0.25, 5.5)) +
  labs(x = "Odds Ratio (log scale)", y = NULL,
       title = "Bayesian Meta-Analysis: PTCy Effect by Comparator",
       subtitle = "Post-Block-9 update — includes RRM and BK virus",
       shape = NULL, colour = NULL, fill = NULL) +
  theme_jama()
```
