---
title: "Comparison 1: PTCy vs CNI+MTX/MMF"
---

```{r}
#| label: setup
#| include: false
source("_common.R")

# Load C1 posteriors
post_os      <- load_post("post_os.rds")
post_m2_os   <- load_post("post_m2_os.rds")
post_nrm     <- load_post("post_nrm.rds")
post_m2_nrm  <- load_post("post_m2_nrm.rds")
post_agvhd   <- load_post("post_agvhd.rds")
post_m1      <- load_post("post_m1.rds")
post_m2      <- load_post("post_m2.rds")
post_m3      <- tryCatch(load_post("post_m3_agvhd.rds"), error = function(e) NULL)
post_bsi     <- load_post("post_bsi.rds")
post_ifi     <- load_post("post_ifi.rds")
post_rrm     <- tryCatch(load_post("post_c1_rrm.rds"), error = function(e) NULL)
post_m2_rrm  <- tryCatch(load_post("post_m2_c1_rrm.rds"), error = function(e) {
  # Fallback: extract from model object
  m <- tryCatch(load_model("m2_c1_rrm.rds"), error = function(e2) NULL)
  if (!is.null(m)) as_draws_df(m) else NULL
})
post_bk      <- tryCatch(load_post("post_c1_bk.rds"), error = function(e) NULL)

# Load analytic datasets (post-Block-9 preferred, fallback to pre)
c1_os    <- load_model("os_brms_clean.rds")
c1_nrm   <- load_model("nrm_brms_clean.rds")
c1_agvhd <- load_model("agvhd_brms_clean.rds")
c1_cmv   <- load_model("cmv_brms_clean.rds")
c1_bsi   <- load_model("bsi_brms_clean.rds")
c1_ifi   <- load_model("ifi_brms_clean.rds")

# Load frequentist models
res_os_freq    <- load_model("res_os_freq.rds")
res_nrm_freq   <- load_model("res_nrm_freq.rds")
res_agvhd_freq <- load_model("res_agvhd_freq.rds")
```

Comparison 1 evaluates PTCy against calcineurin inhibitor–based prophylaxis (CNI+MTX or CNI+MMF) — the most common alternative platform. This comparison draws on the largest evidence base in the meta-analysis.

## Overall Survival {#sec-c1-os}

```{r}
#| label: fig-c1-os-posterior
#| fig-cap: "Posterior density for the PTCy log-OR on overall survival (Comparison 1). M1 (direct effect, navy) and M2 (steroid-adjusted, grey). The vertical dashed line marks the null (log-OR = 0). Shaded regions show 95% credible intervals."
#| fig-height: 4.5

tibble(
  M1 = post_os$b_ptcy_binary,
  M2 = post_m2_os[[ptcy_col(post_m2_os)]]
) |>
  pivot_longer(everything(), names_to = "Model", values_to = "log_or") |>
  ggplot(aes(x = log_or, fill = Model, colour = Model)) +
  geom_density(alpha = 0.35, linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = jama_grey) +
  scale_fill_manual(values = c("M1" = jama_navy, "M2" = jama_steel)) +
  scale_colour_manual(values = c("M1" = jama_navy, "M2" = jama_steel)) +
  labs(x = "log-OR (PTCy vs CNI+MTX/MMF)", y = "Posterior density",
       title = "Overall Survival — Posterior Distributions") +
  theme_jama()
```

**M1 (Direct effect):** OR 0.79 (95% CrI 0.73–0.85), P(OR < 1) = 100%, τ = 0.61. The posterior is entirely below the null, indicating a consistent mortality reduction with PTCy.

**M2 (Steroid-adjusted):** Adjusting for arm-level systemic steroid exposure attenuated the OR from 0.79 to 0.86 (95% CrI 0.77–0.96), a 36% reduction in the log-OR. The residual benefit after adjustment indicates GVHD mediation does not fully explain the OS advantage.

## Non-Relapse Mortality {#sec-c1-nrm}

```{r}
#| label: fig-c1-nrm-posterior
#| fig-cap: "Posterior density for the PTCy log-OR on non-relapse mortality. M2 adjustment abolishes the M1 benefit entirely — the hallmark of complete steroid mediation."
#| fig-height: 4.5

tibble(
  M1 = post_nrm$b_ptcy_binary,
  M2 = post_m2_nrm[[ptcy_col(post_m2_nrm)]]
) |>
  pivot_longer(everything(), names_to = "Model", values_to = "log_or") |>
  ggplot(aes(x = log_or, fill = Model, colour = Model)) +
  geom_density(alpha = 0.35, linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = jama_grey) +
  scale_fill_manual(values = c("M1" = jama_navy, "M2" = jama_steel)) +
  scale_colour_manual(values = c("M1" = jama_navy, "M2" = jama_steel)) +
  labs(x = "log-OR (PTCy vs CNI+MTX/MMF)", y = "Posterior density",
       title = "Non-Relapse Mortality — Posterior Distributions") +
  theme_jama()
```

**M1:** OR 0.88 (95% CrI 0.66–1.18), P(OR < 1) = 83%, τ = 0.72. Directionally favourable but CrI crosses null; k decreased from 15 (pre-Block-9) to 12 after arm reclassification.

**M2:** OR 0.90 (95% CrI 0.59–1.36), P(OR < 1) = 60%. The M1→M2 shift is modest (0.88 → 0.90), with wide CrIs overlapping substantially. Unlike the pre-Block-9 analysis (where M2 crossed null at 1.02, suggesting complete mediation), the current data are **inconclusive regarding steroid mediation** of NRM — the point estimates barely differ, and the wide intervals accommodate both full mediation and no mediation.

::: {.callout-note}
## NRM Mediation — Revised
The pre-Block-9 M1→M2 shift (OR 0.86 → 1.02) was the strongest mediation finding. After arm reclassification and expanded data, the shift is now minimal (0.88 → 0.90) with wide CrIs. The steroid mediation hypothesis for NRM remains plausible but is no longer clearly supported by the Bayesian models alone.
:::

## Mediation Comparison: OS vs NRM {#sec-c1-mediation}

```{r}
#| label: fig-c1-mediation
#| fig-cap: "Comparison of M1→M2 attenuation for OS and NRM. For OS, 71% of the log-OR survives steroid adjustment (partial mediation). For NRM, 0% survives (complete mediation). This contrast defines the causal pathway."
#| fig-height: 5

med_data <- tibble(
  Outcome = rep(c("OS", "NRM"), each = 2),
  Model = rep(c("M1", "M2"), 2),
  or = c(
    median(exp(post_os$b_ptcy_binary)),
    median(exp(post_m2_os[[ptcy_col(post_m2_os)]])),
    median(exp(post_nrm$b_ptcy_binary)),
    median(exp(post_m2_nrm[[ptcy_col(post_m2_nrm)]]))
  ),
  lo = c(
    quantile(exp(post_os$b_ptcy_binary), 0.025),
    quantile(exp(post_m2_os[[ptcy_col(post_m2_os)]]), 0.025),
    quantile(exp(post_nrm$b_ptcy_binary), 0.025),
    quantile(exp(post_m2_nrm[[ptcy_col(post_m2_nrm)]]), 0.025)
  ),
  hi = c(
    quantile(exp(post_os$b_ptcy_binary), 0.975),
    quantile(exp(post_m2_os[[ptcy_col(post_m2_os)]]), 0.975),
    quantile(exp(post_nrm$b_ptcy_binary), 0.975),
    quantile(exp(post_m2_nrm[[ptcy_col(post_m2_nrm)]]), 0.975)
  )
) |>
  mutate(label = sprintf("%.2f [%.2f, %.2f]", or, lo, hi))

ggplot(med_data, aes(x = or, y = interaction(Model, Outcome),
                      colour = Model, shape = Model)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = jama_grey) +
  geom_pointrange(aes(xmin = lo, xmax = hi), size = 0.6, linewidth = 0.6) +
  geom_text(aes(x = hi, label = label), hjust = -0.1, size = 3,
            show.legend = FALSE) +
  scale_colour_manual(values = c("M1" = jama_navy, "M2" = jama_steel)) +
  scale_shape_manual(values = c("M1" = 16, "M2" = 21)) +
  scale_y_discrete(labels = c("M1.NRM" = "NRM — M1", "M2.NRM" = "NRM — M2",
                               "M1.OS" = "OS — M1", "M2.OS" = "OS — M2")) +
  coord_cartesian(xlim = c(0.5, 1.6)) +
  labs(x = "Odds Ratio", y = NULL,
       title = "Steroid Mediation: OS vs NRM",
       subtitle = "M2 = adjusted for arm-level systemic steroid exposure",
       colour = NULL, shape = NULL) +
  theme_jama()
```

## Acute GVHD Grade II–IV {#sec-c1-agvhd}

```{r}
#| label: fig-c1-agvhd
#| fig-cap: "Posterior density for the PTCy effect on aGVHD grade II–IV. This is the largest treatment effect and the mechanistic linchpin of the causal framework."
#| fig-height: 4

tibble(log_or = post_agvhd$b_ptcy_binary) |>
  ggplot(aes(x = log_or)) +
  geom_density(fill = jama_navy, colour = jama_navy, alpha = 0.4, linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = jama_grey) +
  labs(x = "log-OR (PTCy vs CNI+MTX/MMF)", y = "Posterior density",
       title = "aGVHD Grade II\u2013IV — Posterior Distribution",
       subtitle = "k = 28 studies, N = 4,197 · OR 0.67 [0.59\u20130.78]") +
  theme_jama()
```

**M1:** OR 0.67 (95% CrI 0.59–0.78), P(OR < 1) = 100%, τ = 0.68. The largest treatment effect across all outcomes, mechanistically expected given PTCy's mode of action.

## CMV Any-Reactivation {#sec-c1-cmv}

```{r}
#| label: fig-c1-cmv-posterior
#| fig-cap: "Posterior densities for PTCy effect on CMV reactivation. M1 (direct), M2 (steroid-adjusted), and M3 (aGVHD-adjusted). None of the adjustments attenuate the signal — confirming steroid-independent, GVHD-independent harm."
#| fig-height: 4.5

cmv_draws <- tibble(
  M1 = post_m1$b_ptcy_binary,
  `M2 (steroid)` = if ("b_ptcy_binary" %in% names(post_m2)) post_m2$b_ptcy_binary else post_m2$b_eventsn_ptcy_binary
)
if (!is.null(post_m3)) {
  m3_col <- if ("b_ptcy_binary" %in% names(post_m3)) "b_ptcy_binary" else "b_eventsn_ptcy_binary"
  cmv_draws$`M3 (aGVHD)` <- post_m3[[m3_col]]
}
cmv_draws |>
  pivot_longer(everything(), names_to = "Model", values_to = "log_or") |>
  mutate(Model = factor(Model, levels = c("M1", "M2 (steroid)", "M3 (aGVHD)"))) |>
  ggplot(aes(x = log_or, fill = Model, colour = Model)) +
  geom_density(alpha = 0.3, linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = jama_grey) +
  scale_fill_manual(values = c("M1" = jama_navy, "M2 (steroid)" = jama_steel,
                                "M3 (aGVHD)" = jama_light_blue)) +
  scale_colour_manual(values = c("M1" = jama_navy, "M2 (steroid)" = jama_steel,
                                  "M3 (aGVHD)" = jama_light_blue)) +
  labs(x = "log-OR (PTCy vs CNI+MTX/MMF)", y = "Posterior density",
       title = "CMV Reactivation — Posterior Distributions",
       subtitle = "k = 22, N = 3,328 · No attenuation with steroid adjustment") +
  theme_jama()
```

**M1:** OR 1.26 (95% CrI 1.07–1.47), P(OR > 1) > 99%, τ = 0.74. The estimate attenuated from the pre-Block-9 value (1.41) with the expanded corpus, but the CrI still excludes null.

**M2 (steroid-adjusted):** OR 1.25 (no attenuation). The CMV excess is steroid-independent, consistent with direct impairment of CMV-specific T-cell reconstitution rather than a downstream GVHD effect.

## BSI and IFI {#sec-c1-infections}

```{r}
#| label: fig-c1-bsi-ifi
#| fig-cap: "Posterior densities for BSI (any pathogen) and IFI (any). BSI shows a large, consistent excess; IFI shows a probable protective signal with extreme heterogeneity."
#| fig-height: 4.5

tibble(
  BSI = post_bsi$b_ptcy_binary,
  IFI = post_ifi$b_ptcy_binary
) |>
  pivot_longer(everything(), names_to = "Outcome", values_to = "log_or") |>
  ggplot(aes(x = log_or, fill = Outcome, colour = Outcome)) +
  geom_density(alpha = 0.35, linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = jama_grey) +
  scale_fill_manual(values = c("BSI" = jama_navy, "IFI" = jama_light_blue)) +
  scale_colour_manual(values = c("BSI" = jama_navy, "IFI" = jama_light_blue)) +
  labs(x = "log-OR (PTCy vs CNI+MTX/MMF)", y = "Posterior density",
       title = "Infection Outcomes — Posterior Distributions",
       subtitle = "BSI: OR 1.87 [1.33\u20132.62], k = 6  |  IFI: OR 0.43 [0.29\u20130.63], k = 6") +
  theme_jama()
```

**BSI (k = 6):** OR 1.87 (95% CrI 1.33–2.62), τ = 0.96. Elevated risk, attenuated from the pre-Block-9 estimate (2.39) but CrI still excludes null. The frequentist estimate (OR 1.29 [0.61–2.73]) crosses null, reflecting sensitivity to the REML τ² estimate.

**IFI (k = 6):** OR 0.43 (95% CrI 0.29–0.63), P(OR < 1) > 99%. A strong apparent protective signal, substantially stronger than pre-Block-9 (0.73). However, τ = 1.64 indicates extreme heterogeneity and the frequentist CrI (0.21–1.74) crosses null. With k = 6 and this heterogeneity, the estimate should be interpreted cautiously.

## Relapse-Related Mortality {#sec-c1-rrm}

```{r}
#| label: fig-c1-rrm
#| fig-cap: "Posterior density for the PTCy effect on relapse-related mortality (C1). M1 and M2 (steroid-adjusted) are near-identical, indicating no steroid mediation — consistent with a direct graft-versus-leukaemia mechanism."
#| fig-height: 4


if (is.null(post_rrm)) { cat("RRM posteriors not available.") } else {
rrm_draws <- tibble(M1 = post_rrm$b_ptcy_binary)
if (!is.null(post_m2_rrm)) {
  rrm_col <- if ("b_ptcy_binary" %in% names(post_m2_rrm)) "b_ptcy_binary" else "b_eventsn_ptcy_binary"
  rrm_draws$M2 <- post_m2_rrm[[rrm_col]]
}
rrm_draws |>
  pivot_longer(everything(), names_to = "Model", values_to = "log_or") |>
  ggplot(aes(x = log_or, fill = Model, colour = Model)) +
  geom_density(alpha = 0.35, linewidth = 0.6) +
  geom_vline(xintercept = 0, linetype = "dashed", colour = jama_grey) +
  scale_fill_manual(values = c("M1" = jama_navy, "M2" = jama_steel)) +
  scale_colour_manual(values = c("M1" = jama_navy, "M2" = jama_steel)) +
  labs(x = "log-OR (PTCy vs CNI+MTX/MMF)", y = "Posterior density",
       title = "Relapse-Related Mortality — Posterior Distributions",
       subtitle = "k = 38, N = 19,225") +
  theme_jama()
}
```

**M1:** OR 0.84 (95% CrI 0.76–0.93), P(OR < 1) > 99%, k = 38. PTCy is associated with reduced relapse-related mortality, consistent with preserved graft-versus-leukaemia (GVL) activity. This is the first meta-analytic estimate of this outcome.

**M2 (steroid-adjusted):** OR 0.83 (95% CrI 0.73–0.93). No attenuation after steroid adjustment — confirming that the RRM benefit operates through a GVHD-independent pathway, most plausibly GVL preservation. This contrasts with OS (partially mediated) and NRM (uncertain mediation), and helps decompose the residual OS benefit that survives steroid adjustment.

::: {.callout-note}
## RRM Completes the Causal Picture
The RRM finding resolves the question of what drives the residual OS benefit after steroid adjustment. OS M2 retains ~64% of the M1 log-OR; RRM retains ~100%. The portion of the OS advantage not explained by GVHD/steroid reduction is likely attributable to preserved anti-leukaemia immunity.
:::

## BK Virus Reactivation {#sec-c1-bk}

```{r}
#| label: fig-c1-bk
#| fig-cap: "Posterior density for the PTCy effect on BK virus reactivation (C1)."
#| fig-height: 4


if (!is.null(post_bk)) {
  tibble(log_or = post_bk$b_ptcy_binary) |>
    ggplot(aes(x = log_or)) +
    geom_density(fill = jama_navy, colour = jama_navy, alpha = 0.4, linewidth = 0.6) +
    geom_vline(xintercept = 0, linetype = "dashed", colour = jama_grey) +
    labs(x = "log-OR (PTCy vs CNI+MTX/MMF)", y = "Posterior density",
         title = "BK Virus Reactivation — Posterior Distribution",
         subtitle = "k = 11, N = 1,333 · OR 2.48 [1.82\u20133.38]") +
    theme_jama()
}
```

**M1:** OR 2.48 (95% CrI 1.82–3.38), P(OR > 1) > 99%, k = 11, τ = 1.06. The largest infection-outcome effect in the analysis. BK virus reactivation (haemorrhagic cystitis) is substantially more common with PTCy, consistent with cyclophosphamide-induced urothelial damage compounded by impaired BK-specific T-cell surveillance.

::: {.callout-important}
## BK Virus: Dual Mechanism
Unlike CMV and BSI (which operate through T-cell lymphodepletion alone), BK virus reactivation has a **dual mechanism**: (1) direct urothelial toxicity from cyclophosphamide metabolites (acrolein), and (2) impaired BK-specific immune reconstitution. This explains why the BK effect (OR 2.48) is larger than CMV (OR 1.26) and BSI (OR 1.87).
:::

## Summary Forest Plot {#sec-c1-forest}

```{r}
#| label: fig-c1-forest
#| fig-cap: "Summary forest plot for all Comparison 1 outcomes. Filled circles = M1 (direct); open circles = M2 (steroid-adjusted). The null line (OR = 1) is shown as a dashed vertical line."
#| fig-height: 8

c1_summaries <- list(
  or_summary(post_os,    "b_ptcy_binary", "OS", "C1", "M1"),
  or_summary(post_m2_os, ptcy_col(post_m2_os), "OS", "C1", "M2"),
  or_summary(post_nrm,    "b_ptcy_binary", "NRM", "C1", "M1"),
  or_summary(post_m2_nrm, ptcy_col(post_m2_nrm), "NRM", "C1", "M2"),
  or_summary(post_agvhd, "b_ptcy_binary", "aGVHD II\u2013IV", "C1", "M1"),
  or_summary(post_m1,    "b_ptcy_binary", "CMV", "C1", "M1"),
  or_summary(post_m2,    ptcy_col(post_m2), "CMV", "C1", "M2"),
  or_summary(post_bsi,   "b_ptcy_binary", "BSI", "C1", "M1"),
  or_summary(post_ifi,   "b_ptcy_binary", "IFI", "C1", "M1")
)
if (!is.null(post_rrm)) {
  c1_summaries <- c(c1_summaries, list(
    or_summary(post_rrm, "b_ptcy_binary", "RRM", "C1", "M1")
  ))
}
if (!is.null(post_bk)) {
  c1_summaries <- c(c1_summaries, list(
    or_summary(post_bk, "b_ptcy_binary", "BK Virus", "C1", "M1")
  ))
}

outcome_levels <- rev(c("OS", "NRM", "RRM", "aGVHD II\u2013IV", "CMV", "BK Virus", "BSI", "IFI"))

c1_pd <- do.call(rbind, c1_summaries) |>
  as_tibble() |>
  mutate(
    outcome = factor(outcome, levels = outcome_levels),
    label = sprintf("%.2f [%.2f, %.2f]", or, lo, hi),
    y_num = as.numeric(outcome),
    y_pos = y_num + ifelse(model == "M2", -0.15, 0.15)
  )

n_outcomes <- length(unique(c1_pd$outcome))

ggplot(c1_pd, aes(x = or, y = y_pos, shape = model, fill = model, colour = model)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = jama_grey) +
  geom_pointrange(aes(xmin = lo, xmax = hi), size = 0.5, linewidth = 0.5) +
  geom_text(aes(x = hi, label = label), hjust = -0.05, size = 2.6,
            colour = jama_black, show.legend = FALSE, nudge_y = 0.25) +
  scale_x_log10(breaks = c(0.3, 0.5, 0.7, 1, 1.5, 2, 3)) +
  scale_y_continuous(breaks = seq_len(n_outcomes),
                     labels = outcome_levels[seq_len(n_outcomes)]) +
  scale_shape_manual(values = c("M1" = 16, "M2" = 21)) +
  scale_colour_manual(values = c("M1" = jama_navy, "M2" = jama_steel)) +
  scale_fill_manual(values = c("M1" = jama_navy, "M2" = "white")) +
  coord_cartesian(xlim = c(0.25, 5.5)) +
  labs(x = "Odds Ratio (log scale)", y = NULL,
       title = "Comparison 1: PTCy vs CNI+MTX/MMF",
       subtitle = "Bayesian M1/M2 posterior summaries — post-Block-9",
       shape = NULL, colour = NULL, fill = NULL) +
  theme_jama()
```
