---
title: "Risk of Bias"
---

```{r}
#| label: setup
#| include: false
source("_common.R")

rob <- tryCatch(
  read_csv(file.path(extract_dir, "rob.csv"), show_col_types = FALSE),
  error = function(e) NULL
)
```

## Assessment Framework {#sec-rob-framework}

Risk of bias (ROB) was assessed using the ROBINS-I tool (Risk Of Bias In Non-randomised Studies – of Interventions), adapted for comparative observational studies in transplant medicine. Seven domains were evaluated:

1. **Confounding** — baseline differences between PTCy and comparator arms
2. **Selection** — participant selection into study or comparison group
3. **Classification** — misclassification of the intervention (GVHD prophylaxis)
4. **Deviations** — departures from intended interventions
5. **Missing data** — incomplete outcome data
6. **Measurement** — outcome measurement bias
7. **Reporting** — selective reporting of outcomes

Each domain was rated as **Low**, **Moderate**, **Serious**, or **Critical** risk of bias.

## ROB Summary {#sec-rob-summary}

```{r}
#| label: fig-rob-summary
#| fig-cap: "Risk-of-bias summary across all included studies. Most studies carry moderate or serious risk of bias, primarily in the confounding and selection domains — consistent with the observational study design."
#| fig-height: 6

if (!is.null(rob)) {
  domain_labels <- c(
    d1 = "Confounding", d2 = "Selection", d3 = "Classification",
    d4 = "Deviations", d5 = "Missing data", d6 = "Measurement", d7 = "Reporting"
  )

  rob_long <- rob |>
    select(study_id, d1:d7) |>
    pivot_longer(-study_id, names_to = "domain", values_to = "judgment") |>
    mutate(
      domain = domain_labels[domain],
      judgment = factor(judgment,
                        levels = c("Low", "Moderate", "Serious", "Critical", "NR"),
                        labels = c("Low", "Moderate", "Serious", "Critical", "Not reported"))
    ) |>
    filter(!is.na(judgment))

  rob_counts <- rob_long |>
    count(domain, judgment) |>
    group_by(domain) |>
    mutate(pct = n / sum(n) * 100)

  ggplot(rob_counts, aes(x = pct, y = reorder(domain, pct), fill = judgment)) +
    geom_col(width = 0.7) +
    scale_fill_manual(values = c(
      "Low" = "#a3c1d4",
      "Moderate" = jama_light_grey,
      "Serious" = jama_steel,
      "Critical" = jama_dark,
      "Not reported" = "white"
    ), drop = FALSE) +
    labs(x = "Percentage of studies", y = NULL, fill = "Judgment",
         title = "Risk-of-Bias Summary by Domain") +
    theme_jama() +
    theme(legend.position = "bottom")
} else {
  cat("ROB data not found in extraction directory. This page will be populated when rob.csv is available.")
}
```

## Key ROB Considerations by Outcome {#sec-rob-outcomes}

### Overall Survival

| Concern | Assessment |
|---|---|
| **Confounding** | Moderate–Serious. PTCy arms enriched for haploidentical donors (younger, fitter recipients). M2 steroid adjustment partially addresses GVHD-mediated confounding but cannot eliminate all baseline imbalances. |
| **Selection** | Moderate. Registry-based studies reduce individual selection bias but introduce centre-level selection. |
| **Measurement** | Low. OS is an objective endpoint not subject to ascertainment bias. |

### CMV Reactivation

| Concern | Assessment |
|---|---|
| **Confounding** | Serious. Donor CMV serostatus, letermovir prophylaxis, and monitoring intensity vary by arm and era. |
| **Measurement** | Moderate. CMV monitoring frequency may differ between PTCy and comparator arms (awareness bias). More intensive surveillance in PTCy arms could inflate ascertainment. |
| **Reporting** | Low. CMV reactivation is routinely reported in transplant cohorts. |

### BSI

| Concern | Assessment |
|---|---|
| **Missing data** | Very serious (C1). 7 of 13 eligible studies had non-reported event counts, creating substantial attrition bias. |
| **Classification** | Moderate (C2). Outcome definitions varied (any positive culture vs ICU/fatal only). |
| **Confounding** | Serious (C2). Donor-type imbalance in the dominant study (Meyer 2025). |

### IFI

| Concern | Assessment |
|---|---|
| **Confounding** | Serious. Antifungal prophylaxis regimens are heterogeneous and non-randomly distributed. PTCy centres disproportionately use prolonged mold-active azole prophylaxis. |
| **Measurement** | Moderate. IFI diagnosis requires invasive procedures (bronchoscopy/biopsy); investigation intensity may differ between arms. |

## Co-Reviewer Verification {#sec-rob-verification}

::: {.callout-warning}
## Pending
Second-pass ROB verification by BMTD collaborator is pending for all 241 rows in the ROB assessment. This page will be updated upon completion.
:::

## Impact on Interpretation

Risk of bias is a consistent limitation across all outcomes, reflecting the almost exclusively observational evidence base. The mediation models (M2, M3) partially address confounding by adjusting for GVHD-driven steroid exposure, but residual confounding from baseline characteristics (age, disease risk, donor type, conditioning intensity) cannot be excluded.

The most robust findings are those where:

- The effect survives covariate adjustment (OS, CMV in C1)
- The direction is consistent across comparators with different confounding structures (OS, aGVHD)
- Plausible bias would oppose rather than inflate the observed effect (aGVHD)
