---
title: "Sensitivity Analyses"
---

```{r}
#| label: setup
#| include: false
source("_common.R")

# C1 sensitivity models (may not exist in post-Block-9)
post_m1_post2020 <- tryCatch(load_post("post_m1_post2020.rds"), error = function(e) NULL)
post_m1_haplo    <- tryCatch(load_post("post_m1_haplo.rds"), error = function(e) NULL)

# C1 CMV primary
post_m1 <- load_post("post_m1.rds")

# C2 primary and sensitivity
post_c2_os     <- load_post("post_c2_os.rds")
post_c2_agvhd  <- load_post("post_c2_agvhd.rds")
post_c2_cmv    <- load_post("post_c2_cmv.rds")
post_c2_sens_os     <- tryCatch(load_post("post_c2_sens_os.rds"), error = function(e) NULL)
post_c2_sens_agvhd  <- tryCatch(load_post("post_c2_sens_agvhd.rds"), error = function(e) NULL)
post_c2_sens_cmv    <- tryCatch(load_post("post_c2_sens_cmv.rds"), error = function(e) NULL)

# IFI LOO
loo_tbl <- tryCatch(
  readRDS(file.path(model_dir, "loo_tbl.rds")),
  error = function(e) tryCatch(
    readRDS(file.path(model_dir, "loo_ifi.rds")),
    error = function(e2) NULL
  )
)
```

## Comparison 1: CMV Sensitivity Analyses {#sec-sens-c1-cmv}

Two pre-specified sensitivity analyses were conducted for CMV to address era effects and donor-type heterogeneity.

```{r}
#| label: fig-sens-c1-cmv
#| fig-cap: "CMV sensitivity analyses (C1). Both restrictions strengthen the effect, consistent with (a) post-2020 letermovir-era prophylaxis widening the gap, and (b) haploidentical donor settings amplifying PTCy's T-cell depletion effect."
#| fig-height: 5

sens_list <- list(
  or_summary(post_m1, "b_ptcy_binary", "Full corpus", model = "Primary")
)
if (!is.null(post_m1_post2020)) {
  sens_list[[length(sens_list)+1]] <- or_summary(post_m1_post2020, "b_ptcy_binary", "Post-2020", model = "Post-2020")
}
if (!is.null(post_m1_haplo)) {
  sens_list[[length(sens_list)+1]] <- or_summary(post_m1_haplo, "b_ptcy_binary", "Haplo-only", model = "Haplo \u226550%")
}
sens_cmv <- do.call(rbind, sens_list) |>
  as_tibble() |>
  mutate(
    model = factor(model, levels = rev(c("Primary (k=18)", "Post-2020 (k=13)", "Haplo \u226550% (k=11)"))),
    label = sprintf("%.2f [%.2f, %.2f]", or, lo, hi)
  )

ggplot(sens_cmv, aes(x = or, y = model)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = jama_grey) +
  geom_pointrange(aes(xmin = lo, xmax = hi), colour = jama_navy, fill = jama_navy,
                  shape = 16, size = 0.6, linewidth = 0.6) +
  geom_text(aes(x = hi, label = label), hjust = -0.1, size = 3.2, colour = jama_black) +
  coord_cartesian(xlim = c(0.8, 3)) +
  scale_x_log10(breaks = c(1, 1.5, 2, 2.5)) +
  labs(x = "Odds Ratio (log scale)", y = NULL,
       title = "CMV Sensitivity — Comparison 1",
       subtitle = "Restricting to post-2020 or haplo-dominant studies strengthens the signal") +
  theme_jama()
```

| Analysis | k | OR [95% CrI] | Notes |
|---|---|---|---|
| Full corpus | 22 | 1.26 [1.07–1.47] | Attenuated from pre-B9 (1.41) |
| Post-2020 only | — | — | Sensitivity model pre-Block-9 only |
| Haplo ≥50% | — | — | Sensitivity model pre-Block-9 only |

## Comparison 1: IFI Leave-One-Out {#sec-sens-c1-ifi}

```{r}
#| label: fig-sens-ifi-loo
#| fig-cap: "Leave-one-out analysis for IFI (C1). Excluding Yanada 2026 (zero-event structural outlier) collapses τ from 1.05 to 0.38 while barely moving the OR. No other single study produces a comparable influence."
#| fig-height: 5

if (is.data.frame(loo_tbl)) {
  loo_plot <- loo_tbl |>
    mutate(fmt_label = sprintf("%.2f [%.2f, %.2f]", or_median, or_lo, or_hi),
           study = fct_reorder(label, or_median))

  ggplot(loo_plot, aes(x = or_median, y = study)) +
    geom_vline(xintercept = 1, linetype = "dashed", colour = jama_grey) +
    geom_pointrange(aes(xmin = or_lo, xmax = or_hi), colour = jama_navy,
                    size = 0.5, linewidth = 0.5) +
    geom_text(aes(x = or_hi + 0.05, label = sprintf("\u03c4=%.2f", tau_median)),
              hjust = -0.1, size = 2.8, colour = jama_steel) +
    labs(x = "OR (IFI, PTCy vs CNI+MTX/MMF)", y = "Excluded study",
         title = "IFI Leave-One-Out — Comparison 1",
         subtitle = "\u03c4 values annotated; Yanada 2026 drives nearly all heterogeneity") +
    theme_jama()
} else {
  cat("LOO table not available.")
}
```

## Comparison 2: Expanded Arm Inclusion {#sec-sens-c2-arms}

The primary C2 analysis excluded 12 arms (PTCy+ATG combo, mixed, neither). The sensitivity reclassifies PTCy+ATG as PTCy-side and mixed-ATG arms by PTCy co-administration status, adding 6 previously unpaired studies.

```{r}
#| label: fig-sens-c2-arms
#| fig-cap: "Sensitivity analysis — expanded arm inclusion (C2). OS and aGVHD are robust; CMV attenuates notably (OR 0.77 → 0.91) because PTCy+ATG combination arms (reclassified as PTCy-side) carry higher CMV reactivation risk."
#| fig-height: 5

sens_c2_list <- list(
  or_summary(post_c2_os, "b_ptcy_binary", "OS", model = "Primary")
)
if (!is.null(post_c2_sens_os)) {
  sens_c2_list[[length(sens_c2_list)+1]] <- or_summary(post_c2_sens_os, "b_ptcy_binary", "OS", model = "Sensitivity")
}
sens_c2_list[[length(sens_c2_list)+1]] <- or_summary(post_c2_agvhd, "b_ptcy_binary", "aGVHD", model = "Primary")
if (!is.null(post_c2_sens_agvhd)) {
  sens_c2_list[[length(sens_c2_list)+1]] <- or_summary(post_c2_sens_agvhd, "b_ptcy_binary", "aGVHD", model = "Sensitivity")
}
sens_c2_list[[length(sens_c2_list)+1]] <- or_summary(post_c2_cmv, "b_ptcy_binary", "CMV", model = "Primary")
if (!is.null(post_c2_sens_cmv)) {
  sens_c2_list[[length(sens_c2_list)+1]] <- or_summary(post_c2_sens_cmv, "b_ptcy_binary", "CMV", model = "Sensitivity")
}
sens_c2 <- do.call(rbind, sens_c2_list) |>
  as_tibble() |>
  mutate(
    outcome = factor(outcome, levels = c("OS", "aGVHD", "CMV")),
    analysis = ifelse(grepl("Primary", model), "Primary", "Sensitivity"),
    label = sprintf("%.2f [%.2f, %.2f]", or, lo, hi)
  )

sens_c2 <- sens_c2 |>
  mutate(
    y_num = as.numeric(outcome) * 2,
    y_pos = y_num + ifelse(analysis == "Primary", 0.2, -0.2)
  )

ggplot(sens_c2, aes(x = or, y = y_pos, colour = analysis, shape = analysis)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = jama_grey) +
  geom_hline(yintercept = c(3, 5), colour = jama_light_grey, linewidth = 0.2) +
  geom_pointrange(aes(xmin = lo, xmax = hi), size = 0.5, linewidth = 0.5) +
  geom_text(aes(x = hi, label = label), hjust = -0.05, size = 2.8,
            colour = jama_black, show.legend = FALSE, nudge_y = 0.3) +
  scale_colour_manual(values = c("Primary" = jama_navy, "Sensitivity" = jama_steel)) +
  scale_shape_manual(values = c("Primary" = 16, "Sensitivity" = 17)) +
  scale_y_continuous(breaks = c(2, 4, 6), labels = c("OS", "aGVHD", "CMV")) +
  coord_cartesian(xlim = c(0.3, 1.5), ylim = c(0.8, 7.2)) +
  labs(x = "Odds Ratio", y = NULL,
       title = "Comparison 2 \u2014 Expanded Arm Sensitivity",
       subtitle = "Primary (circles) vs sensitivity with expanded arm inclusion (triangles)",
       colour = NULL, shape = NULL) +
  theme_jama()
```

## Comparison 2: Haplo Proportion as Effect Modifier {#sec-sens-c2-haplo}

The proportion of haploidentical donors in the PTCy arm varies substantially across C2 studies (0–100%). We tested whether this modifies the PTCy-vs-ATG treatment effect via frequentist meta-regression with `haplo_pct` as a continuous moderator.

```{r}
#| label: setup-metareg
#| include: false

# Load paired data and join haplo % from arms table
arms_df <- read_csv(file.path(extract_dir, "arms.csv"), show_col_types = FALSE)
studies_tbl <- read_csv(file.path(extract_dir, "studies.csv"), show_col_types = FALSE)

arm_haplo <- arms_df |>
  select(arm_id, donor_haplo_pct) |>
  mutate(haplo_pct = as.numeric(ifelse(donor_haplo_pct %in% c("NR", NA), NA, donor_haplo_pct)))

add_haplo_mod <- function(file) {
  read.csv(file.path(model_dir, file)) |>
    left_join(arm_haplo |> select(arm_id, haplo_pct), by = c("ptcy_arm_id" = "arm_id")) |>
    left_join(studies_tbl |> select(study_id, first_author, pub_year), by = "study_id")
}

mr_dat_os    <- add_haplo_mod("data_c2_os.csv")
mr_dat_cmv   <- add_haplo_mod("data_c2_cmv.csv")
mr_dat_agvhd <- add_haplo_mod("data_c2_agvhd.csv")

mr_es_os    <- escalc(measure = "OR", ai = ptcy_e, n1i = ptcy_n, ci = comp_e, n2i = comp_n, data = mr_dat_os)
mr_es_cmv   <- escalc(measure = "OR", ai = ptcy_e, n1i = ptcy_n, ci = comp_e, n2i = comp_n, data = mr_dat_cmv)
mr_es_agvhd <- escalc(measure = "OR", ai = ptcy_e, n1i = ptcy_n, ci = comp_e, n2i = comp_n, data = mr_dat_agvhd)

fit_mr_os    <- rma(yi, vi, mods = ~ haplo_pct, data = mr_es_os, method = "REML")
fit_mr_cmv   <- rma(yi, vi, mods = ~ haplo_pct, data = mr_es_cmv, method = "REML")
fit_mr_agvhd <- rma(yi, vi, mods = ~ haplo_pct, data = mr_es_agvhd, method = "REML")
```

```{r}
#| label: fig-metareg-haplo
#| fig-cap: "Meta-regression of PTCy-vs-ATG treatment effect on haplo donor proportion. Each bubble is a study (size ∝ precision). The blue line and shaded band show the REML-estimated slope and 95% CI. aGVHD shows a significant positive slope (p = 0.035, R² = 47%): the PTCy advantage over ATG attenuates as haplo proportion increases."
#| fig-height: 4.5
#| fig-width: 11

plot_metareg <- function(es_df, fit, outcome_label) {
  df <- data.frame(
    haplo_pct = es_df$haplo_pct,
    or = exp(es_df$yi),
    wi = 1 / es_df$vi
  )

  newdat <- data.frame(haplo_pct = seq(0, 100, 1))
  pred <- predict(fit, newmods = newdat$haplo_pct)
  pred_df <- data.frame(haplo_pct = newdat$haplo_pct,
                         or = exp(pred$pred),
                         lo = exp(pred$ci.lb),
                         hi = exp(pred$ci.ub))

  ggplot(df, aes(x = haplo_pct, y = or)) +
    geom_hline(yintercept = 1, linetype = "dashed", colour = jama_grey) +
    geom_ribbon(data = pred_df, aes(y = or, ymin = lo, ymax = hi),
                alpha = 0.15, fill = jama_navy) +
    geom_line(data = pred_df, aes(y = or), colour = jama_navy, linewidth = 0.7) +
    geom_point(aes(size = wi), colour = jama_navy, alpha = 0.7) +
    scale_y_log10() +
    scale_size_continuous(range = c(1.5, 6), guide = "none") +
    labs(x = "Haplo donor %", y = "Odds Ratio",
         title = outcome_label,
         subtitle = sprintf("p = %.3f, R\u00B2 = %.0f%%", fit$QMp, fit$R2)) +
    theme_jama(base_size = 10)
}

library(patchwork)
plot_metareg(mr_es_os, fit_mr_os, "OS") +
  plot_metareg(mr_es_cmv, fit_mr_cmv, "CMV") +
  plot_metareg(mr_es_agvhd, fit_mr_agvhd, "aGVHD") +
  plot_annotation(title = "Meta-Regression: Haplo Proportion as Moderator (C2, Primary Dataset)")
```

```{r}
#| label: tbl-metareg-haplo
#| tbl-cap: "Meta-regression results: haplo proportion as continuous moderator of the C2 treatment effect. The aGVHD slope is significant and explains 47% of between-study heterogeneity."

# Predicted ORs for aGVHD at key haplo values
pred_0   <- predict(fit_mr_agvhd, newmods = 0)
pred_50  <- predict(fit_mr_agvhd, newmods = 50)
pred_100 <- predict(fit_mr_agvhd, newmods = 100)

tibble::tribble(
  ~Outcome, ~k, ~`β (per 1% haplo)`, ~`p`, ~`R²`, ~`Predicted OR at 0%`, ~`Predicted OR at 100%`,
  "OS",    fit_mr_os$k,
    sprintf("%.4f", fit_mr_os$beta[2]),
    sprintf("%.3f", fit_mr_os$QMp),
    sprintf("%.0f%%", fit_mr_os$R2),
    NA, NA,
  "CMV",   fit_mr_cmv$k,
    sprintf("%.4f", fit_mr_cmv$beta[2]),
    sprintf("%.3f", fit_mr_cmv$QMp),
    sprintf("%.0f%%", fit_mr_cmv$R2),
    NA, NA,
  "aGVHD", fit_mr_agvhd$k,
    sprintf("%.4f", fit_mr_agvhd$beta[2]),
    sprintf("%.3f", fit_mr_agvhd$QMp),
    sprintf("%.0f%%", fit_mr_agvhd$R2),
    fmt_or(exp(pred_0$pred), exp(pred_0$ci.lb), exp(pred_0$ci.ub)),
    fmt_or(exp(pred_100$pred), exp(pred_100$ci.lb), exp(pred_100$ci.ub))
) |>
  kable(align = c("l","c","c","c","c","c","c")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
```

::: {.callout-note}
## Interpretation

When PTCy is used with **matched donors** (0% haplo), the aGVHD advantage over ATG is large (predicted OR `r fmt_or(exp(pred_0$pred), exp(pred_0$ci.lb), exp(pred_0$ci.ub))`). At **100% haplo**, this advantage is fully attenuated (OR `r fmt_or(exp(pred_100$pred), exp(pred_100$ci.lb), exp(pred_100$ci.ub))`). This is mechanistically coherent: HLA mismatch–driven alloreactivity in haplo transplants partially overwhelms the differential GVHD-suppressive benefit of PTCy relative to ATG. OS shows a suggestive but non-significant trend in the same direction (p = `r sprintf("%.2f", fit_mr_os$QMp)`). CMV is unmoderated by donor type.

This finding was confirmed on both the primary C2 dataset (shown here, k = `r fit_mr_agvhd$k`) and the sensitivity dataset including PTCy+ATG combo arms (k = 14, p = 0.039, R² = 31%).

**Conditioning intensity.** An additive model including both haplo proportion and MAC-predominant conditioning (k = 8, 1 study dropped for missing data) increased R² to 73%. The haplo coefficient strengthened (β = 0.015, p = 0.005) while the MAC term was suggestive but non-significant (β = −0.68, p = 0.14), pointing toward a possible independent MAC-associated enhancement of PTCy's relative GVHD advantage. Importantly, haplo and MAC are not confounded — they pull in opposite directions — and the haplo moderator is robust to conditioning adjustment. A formal interaction term was not estimable at this sample size.
:::

## Publication Bias: RoBMA {#sec-sens-robma}

Robust Bayesian model-averaging (RoBMA) jointly tests for the presence of an effect, heterogeneity, and publication bias by averaging across models with and without each component. P(effect) is the posterior probability that the true effect is non-zero; P(pub bias) is the posterior probability that selection models improve fit; BF(bias) is the corresponding inclusion Bayes factor.

```{r}
#| label: setup-robma
#| include: false

robma_row <- function(file, label) {
  fit <- tryCatch(readRDS(file.path(model_dir, file)), error = function(e) NULL)
  if (is.null(fit)) return(NULL)
  # Extract from stored slots to avoid needing RoBMA loaded
  coefs <- fit$coefficients
  smry  <- fit$summary
  k     <- attr(fit$data, "k_final")
  
  mu_mean <- smry["mu", "Mean"]
  mu_lo   <- smry["mu", "0.025"]
  mu_hi   <- smry["mu", "0.975"]
  p_eff   <- coefs["mu (inclusion)"]
  p_bias  <- coefs["bias (inclusion)"]
  bf_bias <- p_bias / (1 - p_bias)  # convert inclusion prob to BF (prior odds = 1)
  
  data.frame(
    Analysis = label, k = k,
    OR_adj = sprintf("%.2f", exp(mu_mean)),
    CrI = sprintf("[%.2f, %.2f]", exp(mu_lo), exp(mu_hi)),
    P_effect = sprintf("%.1f%%", p_eff * 100),
    P_bias = sprintf("%.1f%%", p_bias * 100),
    BF_bias = ifelse(bf_bias > 1000, ">1000", sprintf("%.1f", bf_bias))
  )
}

robma_tbl <- do.call(rbind, Filter(Negate(is.null), list(
  robma_row("robma_c1_os.rds",    "C1 OS"),
  robma_row("robma_c1_nrm.rds",   "C1 NRM"),
  robma_row("robma_c1_agvhd.rds", "C1 aGVHD"),
  robma_row("robma_c1_cmv.rds",   "C1 CMV"),
  robma_row("robma_c1_bsi.rds",   "C1 BSI"),
  robma_row("robma_c1_ifi.rds",   "C1 IFI"),
  robma_row("robma_c2_os.rds",    "C2 OS"),
  robma_row("robma_c2_agvhd.rds", "C2 aGVHD"),
  robma_row("robma_c2_cmv.rds",   "C2 CMV")
)))
```

```{r}
#| label: tbl-robma
#| tbl-cap: "RoBMA publication bias assessment. P(effect) is the model-averaged posterior probability of a non-null effect after accounting for potential selection. BF(bias) > 3 indicates positive evidence for publication bias."

if (!is.null(robma_tbl)) {
  robma_tbl |>
    kable(row.names = FALSE,
          col.names = c("Analysis", "k", "OR (adj)", "95% CrI",
                        "P(effect)", "P(pub bias)", "BF (bias)"),
          align = c("l","c","c","c","c","c","c")) |>
    kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE)
} else {
  cat("RoBMA models not available.")
}
```

::: {.callout-warning}
## Publication Bias Assessment

RoBMA identifies **strong evidence of publication bias** for C1 OS (BF = 10.9) and C1 aGVHD (BF = 14.1), and **moderate evidence** for C2 OS (BF = 5.2) and C2 aGVHD (BF = 4.1). After bias adjustment, model-averaged posterior probabilities of a non-null effect are below 50% for all outcomes.

This does not mean the effects are definitively absent — it means the current evidence base cannot reliably distinguish true effects from publication-bias artifacts under RoBMA's default skeptical priors (Normal(0, 1) on log-OR, 50% prior mass on the null). The bias evidence is strongest for the two outcomes with the largest apparent effects (C1 OS, C1 aGVHD), consistent with selective reporting amplifying favourable results in non-randomised comparisons. C1 CMV and C1 NRM show no meaningful evidence of bias (BF < 1).
:::

## Bayesian vs Frequentist Concordance {#sec-sens-freq}

```{r}
#| label: tbl-freq-concordance
#| tbl-cap: "Bayesian vs frequentist estimates across both comparisons. Agreement is generally good; OS C2 diverges due to small-study outliers regularised by the Bayesian τ prior."

tibble::tribble(
  ~Outcome, ~Comparison, ~`Bayes OR`, ~`Freq OR`, ~I2, ~Concordance,
  "OS",   "C1", "0.79 [0.73-0.85]", "0.77 [0.66-0.90]", "64%", "Good",
  "OS",   "C2", "0.81 [0.74-0.90]", "0.59 [0.32-1.09]", "95%", "Divergent*",
  "NRM",  "C1", "0.88 [0.66-1.18]", "0.91 [0.68-1.23]", "0%",  "Good",
  "NRM",  "C2", "0.63 [0.30-1.31]", "0.63 [0.30-1.34]", "0%",  "Good (k=3)",
  "aGVHD","C1", "0.67 [0.59-0.78]", "0.61 [0.47-0.79]", "62%", "Good",
  "aGVHD","C2", "0.58 [0.44-0.77]", "0.46 [0.26-0.83]", "73%", "Moderate",
  "CMV",  "C1", "1.26 [1.07-1.47]", "1.25 [0.94-1.66]", "62%", "Good",
  "CMV",  "C2", "0.92 [0.73-1.15]", "0.76 [0.47-1.22]", "71%", "Good",
  "IRM",  "C1", "1.35 [1.16-1.57]", "1.01 [0.70-1.45]", "70%", "Divergent**",
  "IRM",  "C2", "0.90 [0.74-1.09]", "0.81 [0.66-1.00]", "0%",  "Good",
  "cGVHD","C1", "0.33 [0.29-0.36]", "0.44 [0.29-0.67]", "80%", "Good",
  "cGVHD","C2", "0.79 [0.55-1.14]", "0.77 [0.40-1.45]", "60%", "Good"
) |>
  kable(align = c("l","c","c","c","c","c")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) |>
  footnote(symbol = c("C2 OS: Driven by extreme outlier studies; Bayesian tau prior regularises.",
                       "C1 IRM: Bayesian binomial model weights large CIBMTR studies heavily; frequentist RE equalises study contributions."))
```
