---
title: "Sensitivity Analyses"
---

```{r}
#| label: setup
#| include: false
source("_common.R")

# C1 sensitivity models
post_m1_post2020 <- load_post("post_m1_post2020.rds")
post_m1_haplo    <- load_post("post_m1_haplo.rds")

# C1 CMV primary
post_m1 <- load_post("post_m1.rds")

# C2 primary and sensitivity
post_c2_os     <- load_post("post_c2_os.rds")
post_c2_agvhd  <- load_post("post_c2_agvhd.rds")
post_c2_cmv    <- load_post("post_c2_cmv.rds")
post_c2_sens_os     <- load_post("post_c2_sens_os.rds")
post_c2_sens_agvhd  <- load_post("post_c2_sens_agvhd.rds")
post_c2_sens_cmv    <- load_post("post_c2_sens_cmv.rds")

# IFI LOO
loo_tbl <- tryCatch(
  readRDS(file.path(model_dir, "loo_tbl.rds")),
  error = function(e) tryCatch(
    readRDS(file.path(model_dir, "loo_ifi.rds")),
    error = function(e2) NULL
  )
)
```

## Comparison 1: CMV Sensitivity Analyses {#sec-sens-c1-cmv}

Two pre-specified sensitivity analyses were conducted for CMV to address era effects and donor-type heterogeneity.

```{r}
#| label: fig-sens-c1-cmv
#| fig-cap: "CMV sensitivity analyses (C1). Both restrictions strengthen the effect, consistent with (a) post-2020 letermovir-era prophylaxis widening the gap, and (b) haploidentical donor settings amplifying PTCy's T-cell depletion effect."
#| fig-height: 5

sens_cmv <- do.call(rbind, list(
  or_summary(post_m1, "b_ptcy_binary", "Full corpus", model = "Primary (k=18)"),
  or_summary(post_m1_post2020, "b_ptcy_binary", "Post-2020", model = "Post-2020 (k=13)"),
  or_summary(post_m1_haplo, "b_ptcy_binary", "Haplo-only", model = "Haplo \u226550% (k=11)")
)) |>
  as_tibble() |>
  mutate(
    model = factor(model, levels = rev(c("Primary (k=18)", "Post-2020 (k=13)", "Haplo \u226550% (k=11)"))),
    label = sprintf("%.2f [%.2f, %.2f]", or, lo, hi)
  )

ggplot(sens_cmv, aes(x = or, y = model)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = jama_grey) +
  geom_pointrange(aes(xmin = lo, xmax = hi), colour = jama_navy, fill = jama_navy,
                  shape = 16, size = 0.6, linewidth = 0.6) +
  geom_text(aes(x = hi, label = label), hjust = -0.1, size = 3.2, colour = jama_black) +
  coord_cartesian(xlim = c(0.8, 3)) +
  scale_x_log10(breaks = c(1, 1.5, 2, 2.5)) +
  labs(x = "Odds Ratio (log scale)", y = NULL,
       title = "CMV Sensitivity — Comparison 1",
       subtitle = "Restricting to post-2020 or haplo-dominant studies strengthens the signal") +
  theme_jama()
```

| Analysis | k | OR [95% CrI] | P(OR > 1.5) |
|---|---|---|---|
| Full corpus | 18 | 1.41 [1.20–1.65] | 22% |
| Post-2020 only | 13 | 1.80 [1.49–2.18] | 97% |
| Haplo ≥50% | 11 | 1.70 [1.39–2.10] | 89% |

## Comparison 1: IFI Leave-One-Out {#sec-sens-c1-ifi}

```{r}
#| label: fig-sens-ifi-loo
#| fig-cap: "Leave-one-out analysis for IFI (C1). Excluding Yanada 2026 (zero-event structural outlier) collapses τ from 1.05 to 0.38 while barely moving the OR. No other single study produces a comparable influence."
#| fig-height: 5

if (is.data.frame(loo_tbl)) {
  loo_plot <- loo_tbl |>
    mutate(fmt_label = sprintf("%.2f [%.2f, %.2f]", or_median, or_lo, or_hi),
           study = fct_reorder(label, or_median))

  ggplot(loo_plot, aes(x = or_median, y = study)) +
    geom_vline(xintercept = 1, linetype = "dashed", colour = jama_grey) +
    geom_pointrange(aes(xmin = or_lo, xmax = or_hi), colour = jama_navy,
                    size = 0.5, linewidth = 0.5) +
    geom_text(aes(x = or_hi + 0.05, label = sprintf("\u03c4=%.2f", tau_median)),
              hjust = -0.1, size = 2.8, colour = jama_steel) +
    labs(x = "OR (IFI, PTCy vs CNI+MTX/MMF)", y = "Excluded study",
         title = "IFI Leave-One-Out — Comparison 1",
         subtitle = "\u03c4 values annotated; Yanada 2026 drives nearly all heterogeneity") +
    theme_jama()
} else {
  cat("LOO table not available.")
}
```

## Comparison 2: Expanded Arm Inclusion {#sec-sens-c2-arms}

The primary C2 analysis excluded 12 arms (PTCy+ATG combo, mixed, neither). The sensitivity reclassifies PTCy+ATG as PTCy-side and mixed-ATG arms by PTCy co-administration status, adding 6 previously unpaired studies.

```{r}
#| label: fig-sens-c2-arms
#| fig-cap: "Sensitivity analysis — expanded arm inclusion (C2). OS and aGVHD are robust; CMV attenuates notably (OR 0.77 → 0.91) because PTCy+ATG combination arms (reclassified as PTCy-side) carry higher CMV reactivation risk."
#| fig-height: 5

sens_c2 <- do.call(rbind, list(
  or_summary(post_c2_os, "b_ptcy_binary", "OS", model = "Primary (k=10)"),
  or_summary(post_c2_sens_os, "b_ptcy_binary", "OS", model = "Sensitivity (k=11)"),
  or_summary(post_c2_agvhd, "b_ptcy_binary", "aGVHD", model = "Primary (k=6)"),
  or_summary(post_c2_sens_agvhd, "b_ptcy_binary", "aGVHD", model = "Sensitivity (k=8)"),
  or_summary(post_c2_cmv, "b_ptcy_binary", "CMV", model = "Primary (k=12)"),
  or_summary(post_c2_sens_cmv, "b_ptcy_binary", "CMV", model = "Sensitivity (k=14)")
)) |>
  as_tibble() |>
  mutate(
    outcome = factor(outcome, levels = c("OS", "aGVHD", "CMV")),
    analysis = ifelse(grepl("Primary", model), "Primary", "Sensitivity"),
    label = sprintf("%.2f [%.2f, %.2f]", or, lo, hi)
  )

sens_c2 <- sens_c2 |>
  mutate(
    y_num = as.numeric(outcome) * 2,
    y_pos = y_num + ifelse(analysis == "Primary", 0.2, -0.2)
  )

ggplot(sens_c2, aes(x = or, y = y_pos, colour = analysis, shape = analysis)) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = jama_grey) +
  geom_hline(yintercept = c(3, 5), colour = jama_light_grey, linewidth = 0.2) +
  geom_pointrange(aes(xmin = lo, xmax = hi), size = 0.5, linewidth = 0.5) +
  geom_text(aes(x = hi, label = label), hjust = -0.05, size = 2.8,
            colour = jama_black, show.legend = FALSE, nudge_y = 0.3) +
  scale_colour_manual(values = c("Primary" = jama_navy, "Sensitivity" = jama_steel)) +
  scale_shape_manual(values = c("Primary" = 16, "Sensitivity" = 17)) +
  scale_y_continuous(breaks = c(2, 4, 6), labels = c("OS", "aGVHD", "CMV")) +
  coord_cartesian(xlim = c(0.3, 1.5), ylim = c(0.8, 7.2)) +
  labs(x = "Odds Ratio", y = NULL,
       title = "Comparison 2 \u2014 Expanded Arm Sensitivity",
       subtitle = "Primary (circles) vs sensitivity with expanded arm inclusion (triangles)",
       colour = NULL, shape = NULL) +
  theme_jama()
```

## Bayesian vs Frequentist Concordance {#sec-sens-freq}

```{r}
#| label: tbl-freq-concordance
#| tbl-cap: "Bayesian vs frequentist estimates across both comparisons. Agreement is generally good; OS C2 diverges due to small-study outliers regularised by the Bayesian τ prior."

tibble::tribble(
  ~Outcome, ~Comparison, ~`Bayes OR`, ~`Freq OR`, ~I2, ~Concordance,
  "OS",   "C1", "0.77 [0.71-0.83]", "0.75 [0.63-0.89]", "68%", "Good",
  "OS",   "C2", "0.69 [0.61-0.78]", "0.53 [0.29-0.96]", "90%", "Divergent*",
  "NRM",  "C1", "0.86 [0.70-1.08]", "0.78 [0.56-1.09]", "39%", "Good",
  "NRM",  "C2", "0.65 [0.43-0.97]", "0.61 [0.39-0.95]", "0%",  "Good",
  "aGVHD","C1", "0.63 [0.54-0.74]", "0.55 [0.40-0.76]", "63%", "Good",
  "aGVHD","C2", "0.59 [0.42-0.84]", "0.44 [0.22-0.88]", "67%", "Moderate",
  "CMV",  "C1", "1.41 [1.20-1.65]", "1.22 [0.87-1.71]", "69%", "Good",
  "CMV",  "C2", "0.77 [0.60-0.99]", "0.67 [0.41-1.09]", "67%", "Good"
) |>
  kable(align = c("l","c","c","c","c","c")) |>
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = TRUE) |>
  footnote(symbol = "Driven by Dybko 2024 (OR 0.07) and Nykolyszyn 2020 (OR 0.17); Bayesian tau prior regularises these outliers.")
```
